---
title: "CANNEX pain/gut-brain cross-feeding summary"
subtitle: "Core-6 and Core-15 analyses"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float: true
    code_folding: hide
    theme: flatly
params:
  data_root: null
  validation_root: null
  strict_input_hash: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  dpi = 180,
  out.width = "100%"
)

suppressPackageStartupMessages({
  library(DiagrammeR)
  library(digest)
  library(dplyr)
  library(ggplot2)
  library(ggalluvial)
  library(ggnewscale)
  library(htmltools)
  library(jsonlite)
  library(readxl)
  library(scales)
  library(knitr)
  library(kableExtra)
  library(stringr)
  library(tidyr)
})

theme_science <- function() {
  theme_minimal(base_family = "Arial", base_size = 11) +
    theme(
      panel.grid.major.y = element_blank(),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_line(colour = "#E5E8EB", linewidth = 0.35),
      axis.text = element_text(colour = "#303841"),
      axis.title = element_text(colour = "#303841"),
      strip.text = element_text(face = "bold", colour = "#25313C"),
      strip.background = element_rect(fill = "#EFF2F4", colour = NA),
      plot.title = element_text(face = "bold", size = 12.5, colour = "#18232D"),
      plot.title.position = "plot",
      legend.title = element_blank(),
      legend.position = "bottom"
    )
}

manuscript_palette <- list(
  deepskyblue = "#00BFFF",
  lightblue = "#ADD8E6",
  orchid = "#DA70D6",
  mediumorchid = "#BA55D3",
  pink = "#FFC0CB",
  violetred1 = "#FF3E96",
  black = "#111111",
  white = "#FFFFFF",
  lightgrey = "#E8E8EE",
  midgrey = "#73737D"
)
```

```{r input-folder-configuration, include=FALSE}
# ============== RUN THIS FIRST WHEN WORKING CHUNK BY CHUNK =============
# Usually leave these as NULL. If an RStudio folder-selection window is not
# available, enter the appropriate folder between quotation marks.
manual_data_root <- NULL
manual_validation_root <- NULL

nonempty_path <- function(x) {
  is.character(x) && length(x) == 1L && !is.na(x) && nzchar(trimws(x))
}

rmd_parameter <- function(name) {
  if (!exists("params", inherits = TRUE)) return(NULL)
  value <- get("params", inherits = TRUE)[[name]]
  if (is.null(value) || (is.character(value) && !nzchar(value))) NULL else value
}

active_report_directory <- function() {
  input <- tryCatch(knitr::current_input(), error = function(e) "")
  if (nonempty_path(input) && file.exists(input)) {
    return(dirname(normalizePath(input, winslash = "/", mustWork = TRUE)))
  }

  if (interactive() && requireNamespace("rstudioapi", quietly = TRUE) &&
      rstudioapi::isAvailable()) {
    source_path <- tryCatch(
      rstudioapi::getSourceEditorContext()$path,
      error = function(e) ""
    )
    if (nonempty_path(source_path) && file.exists(source_path)) {
      return(dirname(normalizePath(
        source_path, winslash = "/", mustWork = TRUE
      )))
    }
  }

  normalizePath(getwd(), winslash = "/", mustWork = TRUE)
}

folder_contains <- function(root, required_files) {
  nonempty_path(root) && dir.exists(root) &&
    all(file.exists(file.path(root, required_files)))
}

select_folder_in_rstudio <- function(caption, starting_directory) {
  if (!interactive() || !requireNamespace("rstudioapi", quietly = TRUE) ||
      !rstudioapi::isAvailable()) {
    return(NULL)
  }
  selected <- tryCatch(
    rstudioapi::selectDirectory(
      caption = caption, label = "Use this folder", path = starting_directory
    ),
    error = function(e) NULL
  )
  if (nonempty_path(selected)) selected else NULL
}

resolve_input_folder <- function(
    label, required_files, manual_value, parameter_name, environment_name,
    automatic_candidates, starting_directory) {
  candidates <- c(
    manual_value,
    rmd_parameter(parameter_name),
    Sys.getenv(environment_name, unset = ""),
    automatic_candidates
  )
  candidates <- unique(candidates[vapply(candidates, nonempty_path, logical(1))])

  for (candidate in candidates) {
    if (folder_contains(candidate, required_files)) {
      return(normalizePath(candidate, winslash = "/", mustWork = TRUE))
    }
  }

  selected <- select_folder_in_rstudio(
    paste0("Select the ", label, " folder"), starting_directory
  )
  if (folder_contains(selected, required_files)) {
    return(normalizePath(selected, winslash = "/", mustWork = TRUE))
  }

  stop(
    "Could not locate the ", label, " folder. Run this configuration chunk ",
    "again and select the folder containing: ",
    paste(required_files, collapse = ", "), ". If no folder-selection window ",
    "appears, set the corresponding manual_*_root value in this chunk."
  )
}

report_directory <- active_report_directory()

data_required_files <- c(
  file.path(
    "crossfeeding_analysis_6", "batch1_pain_gutbrain_results.xlsx"
  ),
  file.path(
    "crossfeeding_analysis_6", "models_qc_passed", "model_manifest.tsv"
  ),
  file.path(
    "crossfeeding_analysis_6", "batch1_single_root_literature_evidence.tsv"
  ),
  file.path(
    "crossfeeding_analysis_15", "batch1_species_root_posthoc.json"
  ),
  file.path(
    "crossfeeding_analysis_15", "batch1_pain_gutbrain_results.xlsx"
  ),
  file.path(
    "crossfeeding_analysis_15", "batch1_single_root_literature_evidence.tsv"
  )
)
validation_required_files <- c(
  "VALIDATION_INDEX.tsv", "validation_summary.csv"
)

resolved_data_root <- resolve_input_folder(
  label = "CANNEX gem_screening",
  required_files = data_required_files,
  manual_value = manual_data_root,
  parameter_name = "data_root",
  environment_name = "CROSSFEEDING_DATA_ROOT",
  automatic_candidates = c(report_directory, getwd()),
  starting_directory = report_directory
)

resolved_validation_root <- resolve_input_folder(
  label = "Khoi_prototype/validation",
  required_files = validation_required_files,
  manual_value = manual_validation_root,
  parameter_name = "validation_root",
  environment_name = "CROSSFEEDING_VALIDATION_ROOT",
  automatic_candidates = c(
    resolved_data_root,
    file.path(report_directory, "validation"),
    file.path(resolved_data_root, "validation"),
    getwd()
  ),
  starting_directory = report_directory
)

crossfeeding_input_paths <- list(
  data_root = resolved_data_root,
  validation_root = resolved_validation_root
)
crossfeeding_strict_input_hash <- rmd_parameter("strict_input_hash")
if (is.null(crossfeeding_strict_input_hash)) {
  crossfeeding_strict_input_hash <- TRUE
}

# Retain the selections for later runs in the same RStudio session.
Sys.setenv(
  CROSSFEEDING_DATA_ROOT = crossfeeding_input_paths$data_root,
  CROSSFEEDING_VALIDATION_ROOT = crossfeeding_input_paths$validation_root
)

cat(
  "Input folders ready.\n",
  "CANNEX data: ", crossfeeding_input_paths$data_root, "\n",
  "Validation data: ", crossfeeding_input_paths$validation_root, "\n",
  sep = ""
)
# ======================================================================
```

## Interpretation boundary

The core Python workflow identifies **GEM-supported potential cross-feeding relationships**. It does not establish realized metabolite exchange in vivo. Core-6 and Core-15 are displayed separately because the taxon sets overlap and therefore must not be summed.

For the literature layer, donor-to-precursor evidence is now mandatory. Two
mutually exclusive qualifying classes are used: **Case 1** means that the
donor-to-exact-precursor reaction is supported but the recipient-to-target
reaction is not; **Case 2** means that both reactions are supported at
exact-strain or same-species resolution. Recipient-to-target evidence alone is
not classified as literature support for a cross-feeding relationship.

## Methods

### CANNEX taxon sets and 16S refinement

Two previously locked CANNEX taxon sets were analyzed independently. The
Core-6 analysis began with six ASVs and retained four ASVs for GEM analysis.
The Core-15 analysis began with 15 ASVs, retained 13 genus- or species-level
assignments for sequence assessment, and retained seven ASVs after 16S
refinement. Candidate AGORA2 taxa were evaluated using only 16S rRNA reference
sequences; whole-genome sequences were not used. The complete 253-nt ASV V4
sequence was aligned to candidate reference sequences, and high-confidence
mapping required complete V4 coverage and at least 99% nucleotide identity.
Only species tied at the highest identity for an ASV were carried forward.
Because this amplicon region cannot establish strain identity, all available
GEMs within a retained species were treated as a strain panel rather than as
proof that the study ASV represented a specific strain.

### GEM selection and quality control

Candidate models were obtained from the locally archived AGORA2 v2.01
collection. Forty GEMs were considered for Core-6 and 50 GEMs were selected
after 16S refinement for Core-15. Each source SBML file was required to parse
as well-formed XML without alteration. Files declaring UTF-8 but containing an
incompatible Latin-1 byte were excluded rather than repaired. This produced 30
raw-XML-QC-passing Core-6 GEMs and 37 Core-15 GEMs; subsequent Core-6 sequence
refinement yielded its final 29-GEM panel. Model manifests recorded species,
strain, source filename, model identifier, and SHA-256 checksum so that the
exact input panels and species denominators were fixed before screening.

### Target selection and cross-feeding screen

Seven pain- or gut-brain-axis-related metabolites were selected before GEM
screening: GABA, butyrate, histamine, deoxycholate, hydrogen sulfide,
tryptamine, and indole-3-propionate. Expected precursors, donors, and recipients
were not supplied to the discovery algorithm. The unchanged target-focused
Python workflow was run with Python 3.11, MeneTools 3.4.0, the AGORA2 HighFiber
diet as study context, anaerobic target-withheld conditions, and
`PYTHONHASHSEED=0` for deterministic search ordering.

For each selected target, the workflow sequentially required: (i) presence of
the target identifier; (ii) a target exchange reaction permitting export;
(iii) directionally valid outward transport from cytosol to extracellular
space, including a cytosol-to-periplasm-to-extracellular chain when represented;
(iv) at least one target-withheld intracellular production path with strictly
positive net target formation; (v) exchange import and directionally valid
inward transport for each nominated extracellular precursor; and (vi)
intracellular precursor formation, outward transport, and export by a
different-species donor while withholding that precursor. MeneTools paths were
qualitative, cardinality-minimal reaction sets rather than flux-, yield-, or
growth-optimal solutions. Up to 25 representative alternative paths were
retained per recipient-model/target pair, together with their complete
non-currency extracellular root sets.

### Post-hoc aggregation and evidence classification

Saved model-level results were audited without rerunning or modifying the core
workflow. Recipient paths were rechecked for positive net target formation,
and retained donor paths were rechecked for positive net precursor production,
outward transport, and exchange export. Results were collapsed using the
directional key target + donor species + focal precursor + recipient species,
while retaining the exact supporting and tested GEM counts (`n/N`). A
relationship was classified as single-precursor when at least one retained
recipient path used only the focal extracellular precursor as its non-currency
root; otherwise it was classified as multi-precursor. For every displayed
multi-precursor Case-2 route, the same recipient GEM was additionally required
to contain the exact production path and to support exchange import and inward
transport of every root in the displayed input set.

Primary-literature assessment was performed after fixing the GEM-derived
relationships. Donor evidence required exact-precursor production or release
at exact-strain or same-species resolution; genus-level evidence, gene presence
alone, and unresolved stereochemistry did not qualify. `Case 1` denoted
donor-to-precursor support without qualifying recipient-to-target support, and
`Case 2` denoted support for both terminal reactions. For the simplified
Sankey display, `Yes` combines Cases 1 and 2 because both contain qualifying
donor-to-exact-precursor evidence; `No` denotes relationships for which no such
donor evidence was identified. The latter is an evidence-gap classification,
not evidence that the modeled relationship is biologically impossible.

### Pipeline validation and visualization

Eight literature-defined validation cases were assessed using the saved
workflow outputs. Two matched triads tested a positive and negative branch
while holding either the recipient or donor constant; three additional cases
served as independent positive anchors. The remaining cases were retained as
partial or informative non-pass examples to expose current GEM and workflow
boundaries. Figures were generated in R using `ggplot2`, `ggalluvial`, and
`DiagrammeR`. Network nodes represent potential donors, extracellular
precursors or complete input sets, potential recipients, and target
metabolites. Edge labels report `n/N` GEM support. These analyses describe
structural metabolic potential and do not establish secretion quantity,
simultaneous community flux, recipient growth benefit, in vivo transfer, or
causality for pain.

## Results

### Pipeline validation controls

Eight literature-defined cases were used to test whether the workflow retains
expected relationships and excludes expected null relationships. Cases 01 and
02 are matched triad controls: each contains a positive branch and a negative
branch while holding either the recipient or donor constant. Cases 03--05 are
independent positive anchors. Case 06 is a partial pathway match, whereas Cases
07--08 are informative non-passes in which target-production potential was
recovered but the literature-described donor edge was not reproduced. These
last three cases are shown to define the present workflow boundary and are not
counted as formal validation passes.

```{r validation-data-preparation, include=FALSE}
if (!exists("crossfeeding_input_paths", inherits = TRUE)) {
  stop(
    "Run the 'input-folder-configuration' chunk before this chunk."
  )
}
validation_root <- crossfeeding_input_paths$validation_root

validation_inputs <- c(
  validation_index = file.path(validation_root, "VALIDATION_INDEX.tsv"),
  validation_summary = file.path(validation_root, "validation_summary.csv")
)

if (!all(file.exists(validation_inputs))) {
  stop(
    "Missing locked validation input(s): ",
    paste(names(validation_inputs)[!file.exists(validation_inputs)], collapse = ", "),
    ". Run the 'input-folder-configuration' chunk again and select the ",
    "Khoi_prototype/validation directory."
  )
}

expected_validation_sha256 <- c(
  validation_index = "1fd302fedec934b5bd84d96384ae06e967f27973fd9b5e88d2ed3159f591a816",
  validation_summary = "bec222a7206efa1d5be9c11158aeb85d692b5bcd2b0b6ee4ad7667ad9df1d69a"
)
observed_validation_sha256 <- vapply(
  validation_inputs,
  function(path) digest(file = path, algo = "sha256", serialize = FALSE),
  character(1)
)
if (isTRUE(crossfeeding_strict_input_hash) &&
    !identical(observed_validation_sha256, expected_validation_sha256)) {
  changed <- names(observed_validation_sha256)[
    observed_validation_sha256 != expected_validation_sha256
  ]
  stop("Locked validation input changed: ", paste(changed, collapse = ", "))
}

validation_index <- read.delim(
  validation_inputs[["validation_index"]],
  check.names = FALSE, stringsAsFactors = FALSE
) |>
  mutate(
    display_case = paste0("Case ", str_sub(case_id, 1, 2)),
    outcome = case_when(
      str_starts(status, "Passed:") ~ "Formal pass",
      str_starts(status, "Retained partial") ~ "Partial match",
      TRUE ~ "Informative non-pass"
    )
  )

validation_summary <- read.csv(
  validation_inputs[["validation_summary"]],
  check.names = FALSE, stringsAsFactors = FALSE
)

if (nrow(validation_index) != 8L || nrow(validation_summary) != 10L) {
  stop("Expected 8 validation cases and 10 case-branch summary rows.")
}
if (!identical(
  as.integer(table(factor(
    validation_index$outcome,
    levels = c("Formal pass", "Partial match", "Informative non-pass")
  ))),
  c(5L, 1L, 2L)
)) {
  stop("The locked case-level validation outcome counts have changed.")
}

validation_outcomes <- validation_index |>
  count(outcome, name = "cases") |>
  mutate(
    outcome = factor(
      outcome,
      levels = c("Formal pass", "Partial match", "Informative non-pass")
    )
  )

validation_html_label <- function(x, width = 28L) {
  lines <- strwrap(x, width = width)
  escaped <- vapply(lines, function(z) as.character(htmlEscape(z)), character(1))
  paste0("<", paste(escaped, collapse = "<BR/>"), ">")
}

network_with_downloads <- function(widget, file_stub, png_scale = 3) {
  # Downloads are created only after a reader clicks a button; knitting the
  # report does not write any figure files.
  element_id <- paste0(
    "network_", gsub("[^A-Za-z0-9_]+", "_", file_stub)
  )
  controls_id <- paste0(element_id, "_download_controls")
  widget$elementId <- element_id

  button_style <- paste(
    "background:#FFFFFF", "border:1px solid #111111",
    "border-radius:4px", "color:#111111", "cursor:pointer",
    "font-family:Arial,sans-serif", "font-size:12px",
    "margin-right:6px", "padding:5px 10px", sep = ";"
  )
  controls <- tags$div(
    id = controls_id,
    style = paste0("margin:6px 0 14px 0;text-align:left;"),
    tags$button(
      type = "button", `data-network-format` = "svg",
      style = button_style, "Download SVG"
    ),
    tags$button(
      type = "button", `data-network-format` = "png",
      style = button_style,
      paste0("Download PNG (", png_scale, "x)")
    )
  )

  javascript <- paste0(
    "(function(){",
    "const targetId=", toJSON(element_id, auto_unbox = TRUE), ";",
    "const controlsId=", toJSON(controls_id, auto_unbox = TRUE), ";",
    "const fileBase=", toJSON(file_stub, auto_unbox = TRUE), ";",
    "const pngScale=", as.numeric(png_scale), ";",
    "const controls=document.getElementById(controlsId);",
    "if(!controls||controls.dataset.downloadReady==='1'){return;}",
    "controls.dataset.downloadReady='1';",
    "function downloadBlob(blob,name){",
    "const url=URL.createObjectURL(blob);",
    "const link=document.createElement('a');",
    "link.href=url;link.download=name;document.body.appendChild(link);",
    "link.click();link.remove();",
    "window.setTimeout(function(){URL.revokeObjectURL(url);},1000);",
    "}",
    "function prepareSvg(svg){",
    "const clone=svg.cloneNode(true);",
    "clone.setAttribute('xmlns','http://www.w3.org/2000/svg');",
    "clone.setAttribute('xmlns:xlink','http://www.w3.org/1999/xlink');",
    "const rect=svg.getBoundingClientRect();",
    "const vb=(svg.viewBox&&svg.viewBox.baseVal)?svg.viewBox.baseVal:null;",
    "const width=Math.max(1,Math.round(rect.width||(vb&&vb.width)||1200));",
    "const height=Math.max(1,Math.round(rect.height||(vb&&vb.height)||700));",
    "clone.setAttribute('width',width);clone.setAttribute('height',height);",
    "const ns='http://www.w3.org/2000/svg';",
    "const bg=document.createElementNS(ns,'rect');",
    "bg.setAttribute('x',vb?vb.x:0);bg.setAttribute('y',vb?vb.y:0);",
    "bg.setAttribute('width',vb?vb.width:width);",
    "bg.setAttribute('height',vb?vb.height:height);",
    "bg.setAttribute('fill','#FFFFFF');clone.insertBefore(bg,clone.firstChild);",
    "return {xml:'<?xml version=\"1.0\" encoding=\"UTF-8\"?>'+",
    "new XMLSerializer().serializeToString(clone),width:width,height:height};",
    "}",
    "controls.addEventListener('click',function(event){",
    "const button=event.target.closest('button[data-network-format]');",
    "if(!button){return;}",
    "const target=document.getElementById(targetId);",
    "const svg=target?target.querySelector('svg'):null;",
    "if(!svg){window.alert('The network is still rendering. Please try again.');return;}",
    "const prepared=prepareSvg(svg);",
    "if(button.dataset.networkFormat==='svg'){",
    "downloadBlob(new Blob([prepared.xml],",
    "{type:'image/svg+xml;charset=utf-8'}),fileBase+'.svg');return;",
    "}",
    "const svgBlob=new Blob([prepared.xml],{type:'image/svg+xml;charset=utf-8'});",
    "const svgUrl=URL.createObjectURL(svgBlob);const image=new Image();",
    "image.onload=function(){",
    "const canvas=document.createElement('canvas');",
    "canvas.width=Math.ceil(prepared.width*pngScale);",
    "canvas.height=Math.ceil(prepared.height*pngScale);",
    "const context=canvas.getContext('2d');",
    "context.setTransform(pngScale,0,0,pngScale,0,0);",
    "context.drawImage(image,0,0,prepared.width,prepared.height);",
    "URL.revokeObjectURL(svgUrl);",
    "canvas.toBlob(function(blob){",
    "if(blob){downloadBlob(blob,fileBase+'.png');}",
    "},'image/png',1);",
    "};",
    "image.onerror=function(){URL.revokeObjectURL(svgUrl);",
    "window.alert('PNG creation failed; please use the SVG download.');};",
    "image.src=svgUrl;",
    "});",
    "})();"
  )

  htmlwidgets::appendContent(
    widget,
    controls,
    tags$script(HTML(javascript))
  )
}

validation_case_specs <- list(
  list(
    case = "Case 01", title = "Two donors; one fixed recipient",
    outcome = "Formal pass",
    nodes = data.frame(
      id = c("dp", "dn", "p", "r", "t"),
      label = c(
        "Bifidobacterium adolescentis L2-32",
        "Paraprevotella xylaniphila YIT 11841",
        "L-lactate", "Eubacterium hallii L2-7", "Butyrate"
      ),
      role = c("donor", "negative", "precursor", "recipient", "target")
    ),
    edges = data.frame(
      from = c("dp", "p", "r", "dn"),
      to = c("p", "r", "t", "p"),
      state = c("retained", "retained", "retained", "confirmed_absent"),
      label = c("retained", "", "", "negative donor excluded")
    )
  ),
  list(
    case = "Case 02", title = "One donor; two recipients",
    outcome = "Formal pass",
    nodes = data.frame(
      id = c("d", "p", "rp", "rn", "t"),
      label = c(
        "Bifidobacterium adolescentis L2-32", "L-lactate",
        "Eubacterium hallii L2-7",
        "Phascolarctobacterium succinatutens YIT 12067", "Butyrate"
      ),
      role = c("donor", "precursor", "recipient", "negative", "target")
    ),
    edges = data.frame(
      from = c("d", "p", "rp", "p"),
      to = c("p", "rp", "t", "rn"),
      state = c("retained", "retained", "retained", "confirmed_absent"),
      label = c("", "retained", "", "negative recipient excluded")
    )
  ),
  list(
    case = "Case 03", title = "Succinate to propionate",
    outcome = "Formal pass",
    nodes = data.frame(
      id = c("d", "p", "r", "t"),
      label = c(
        "Paraprevotella xylaniphila YIT 11841", "Succinate",
        "Phascolarctobacterium succinatutens YIT 12067", "Propionate"
      ),
      role = c("donor", "precursor", "recipient", "target")
    ),
    edges = data.frame(
      from = c("d", "p", "r"), to = c("p", "r", "t"),
      state = rep("retained", 3), label = c("", "retained", "")
    )
  ),
  list(
    case = "Case 04", title = "Acetate to butyrate",
    outcome = "Formal pass",
    nodes = data.frame(
      id = c("d", "p", "r", "t"),
      label = c(
        "Bifidobacterium longum NCC2705", "Acetate",
        "Eubacterium rectale ATCC 33656", "Butyrate"
      ),
      role = c("donor", "precursor", "recipient", "target")
    ),
    edges = data.frame(
      from = c("d", "p", "r"), to = c("p", "r", "t"),
      state = rep("retained", 3), label = c("", "retained", "")
    )
  ),
  list(
    case = "Case 05", title = "L-lactate to hydrogen sulfide",
    outcome = "Formal pass",
    nodes = data.frame(
      id = c("d", "p", "r", "t"),
      label = c(
        "Bifidobacterium adolescentis L2-32", "L-lactate",
        "Desulfovibrio piger ATCC 29098 (DSM 749)", "Hydrogen sulfide"
      ),
      role = c("donor", "precursor", "recipient", "target")
    ),
    edges = data.frame(
      from = c("d", "p", "r"), to = c("p", "r", "t"),
      state = rep("retained", 3), label = c("", "retained", "")
    )
  ),
  list(
    case = "Case 06", title = "Putrescine pathway",
    outcome = "Partial match",
    nodes = data.frame(
      id = c("d", "p", "a", "r", "t"),
      label = c(
        "Streptococcus gordonii CH1", "Ornithine", "L-arginine",
        "Fusobacterium nucleatum ATCC 25586", "Putrescine"
      ),
      role = c("donor", "precursor", "alternative", "recipient", "target")
    ),
    edges = data.frame(
      from = c("d", "p", "d", "a", "r"),
      to = c("p", "r", "a", "r", "t"),
      state = c("not_recovered", "context", "partial", "partial", "partial"),
      label = c(
        "literature edge not recovered", "literature route",
        "GEM-supported alternative", "alternative route",
        "production recovered"
      )
    )
  ),
  list(
    case = "Case 07", title = "Indole-3-lactate to indole-3-propionate",
    outcome = "Informative non-pass",
    nodes = data.frame(
      id = c("d", "p", "r", "t"),
      label = c(
        "Lactobacillus reuteri I5007", "Indole-3-lactate",
        "Clostridium sporogenes ATCC 15579", "Indole-3-propionate"
      ),
      role = c("donor", "precursor", "recipient", "target")
    ),
    edges = data.frame(
      from = c("d", "p", "r"), to = c("p", "r", "t"),
      state = c("not_recovered", "context", "partial"),
      label = c(
        "donor GEM lacks metabolite", "literature route",
        "production recovered"
      )
    )
  ),
  list(
    case = "Case 08", title = "Cholate to deoxycholate",
    outcome = "Informative non-pass",
    nodes = data.frame(
      id = c("d", "p", "r", "t"),
      label = c(
        "Bacteroides thetaiotaomicron VPI-5482", "Cholate",
        "Clostridium scindens ATCC 35704", "Deoxycholate"
      ),
      role = c("donor", "precursor", "recipient", "target")
    ),
    edges = data.frame(
      from = c("d", "p", "r"), to = c("p", "r", "t"),
      state = c("not_recovered", "partial", "partial"),
      label = c(
        "donor production not retained", "recipient precursor found",
        "production recovered"
      )
    )
  )
)

make_validation_composite <- function(specs, style) {
  # Graphviz fills this packed one-column array bottom-to-top. Reverse each
  # pair so the lower-numbered case is displayed above the next case.
  if (length(specs) == 2L) {
    specs <- rev(specs)
  }
  role_styles <- list(
    donor = c(
      style$donor_fill, style$donor_border, style$donor_font_colour
    ),
    negative = c(
      style$negative_fill, style$negative_border, style$negative_font_colour
    ),
    precursor = c(
      style$precursor_fill, style$precursor_border,
      style$precursor_font_colour
    ),
    alternative = c(
      style$alternative_fill, style$alternative_border,
      style$alternative_font_colour
    ),
    recipient = c(
      style$recipient_fill, style$recipient_border,
      style$recipient_font_colour
    ),
    target = c(
      style$target_fill, style$target_border, style$target_font_colour
    )
  )
  edge_styles <- list(
    retained = list(
      colour = style$retained_edge_colour, line = "solid",
      head = style$arrow_head, width = style$retained_edge_width
    ),
    confirmed_absent = list(
      colour = style$confirmed_absent_edge_colour, line = "dashed",
      head = "tee", width = style$absent_edge_width
    ),
    partial = list(
      colour = style$partial_edge_colour, line = "solid",
      head = style$arrow_head, width = style$partial_edge_width
    ),
    not_recovered = list(
      colour = style$not_recovered_edge_colour, line = "dashed",
      head = "tee", width = style$missing_edge_width
    ),
    context = list(
      colour = style$context_edge_colour, line = "dashed",
      head = style$arrow_head, width = style$context_edge_width
    )
  )

  cluster_blocks <- vapply(seq_along(specs), function(case_index) {
    spec <- specs[[case_index]]
    prefix <- paste0("case", str_pad(case_index, 2, pad = "0"), "_")

    node_lines <- vapply(seq_len(nrow(spec$nodes)), function(i) {
      node <- spec$nodes[i, ]
      colours <- role_styles[[node$role]]
      shape <- if (node$role == "precursor") {
        style$precursor_shape
      } else if (node$role == "target") {
        style$target_shape
      } else {
        style$node_shape
      }
      sprintf(
        '%s%s [label=%s, shape="%s", fillcolor="%s", color="%s", fontcolor="%s", penwidth=%.2f];',
        prefix, node$id,
        validation_html_label(node$label, style$label_wrap_width),
        shape, colours[1], colours[2], colours[3], style$node_border_width
      )
    }, character(1))

    edge_lines <- vapply(seq_len(nrow(spec$edges)), function(i) {
      edge <- spec$edges[i, ]
      edge_style <- edge_styles[[edge$state]]
      label_attribute <- if (nzchar(edge$label)) {
        paste0(
          ", label=",
          validation_html_label(edge$label, style$edge_label_wrap_width)
        )
      } else {
        ""
      }
      sprintf(
        '%s%s -> %s%s [color="%s", fontcolor="%s", fontsize=%.1f, style="%s", penwidth=%.2f, arrowhead="%s", arrowsize=%.2f%s];',
        prefix, edge$from, prefix, edge$to,
        edge_style$colour, edge_style$colour,
        style$edge_label_size, edge_style$line, edge_style$width,
        edge_style$head, style$arrow_size, label_attribute
      )
    }, character(1))

    case_colour <- style$case_border
    case_badge <- sprintf(
      '%sbadge [label=%s, shape="box", style="rounded,filled", fillcolor="%s", color="%s", fontcolor="%s", fontsize=%.1f, penwidth=1.1, margin=0.07];',
      prefix,
      validation_html_label(spec$case, 12L),
      style$case_fill, style$case_border, style$case_font_colour,
      style$case_title_size
    )
    badge_anchor <- sprintf(
      '%sbadge -> %s%s [style="invis", weight=80];',
      prefix, prefix, spec$nodes$id[1]
    )
    case_label <- paste0(
      spec$case, " — ", spec$title, " [", spec$outcome, "]"
    )
    paste0(
      "subgraph cluster_", prefix, " {\n",
      "label=", validation_html_label(case_label, style$case_title_wrap_width),
      "; labelloc=\"t\"; labeljust=\"l\"; color=\"", case_colour,
      "\"; fontcolor=\"", case_colour,
      "\"; fontsize=", style$case_title_size,
      "; penwidth=", style$cluster_border_width,
      "; style=\"rounded\"; margin=", style$cluster_margin, ";\n",
      case_badge, "\n",
      paste(node_lines, collapse = "\n"), "\n",
      paste(edge_lines, collapse = "\n"), "\n",
      badge_anchor, "\n}"
    )
  }, character(1))

  dot <- paste0(
    "digraph G {\n",
    "graph [rankdir=LR, bgcolor=\"", style$background_colour,
    "\", pad=", style$graph_padding,
    ", pack=true, packmode=\"array_u2\", newrank=true",
    ", nodesep=", style$node_separation,
    ", ranksep=", style$rank_separation, "];\n",
    "node [style=\"rounded,filled\", fontname=\"", style$font_name,
    "\", fontsize=", style$node_font_size,
    ", width=", style$node_width, ", height=", style$node_height,
    ", margin=", style$node_padding, "];\n",
    "edge [fontname=\"", style$font_name, "\"];\n",
    paste(cluster_blocks, collapse = "\n"), "\n}"
  )

  grViz(dot, width = style$widget_width, height = style$pair_height_px)
}
```

The overview is case-level, so the positive and negative branches in Cases 01
and 02 each contribute to one formal validation case rather than being counted
as separate cases.

```{r validation-control-parameters, include=FALSE}
# ================= USER-EDITABLE: VALIDATION CONTROLS ==================
validation_style <- list(
  # Status overview colours
  outcome_colours = c(
    "Formal pass" = "#00BFFF",
    "Partial match" = "#BA55D3",
    "Informative non-pass" = "#FF3E96"
  ),

  # Status-overview bar plot
  outcome_figure_width = 10, outcome_figure_height = 3.5,
  outcome_output_width = "100%",
  outcome_bar_width = 0.64,
  outcome_label_colour = "#111111", outcome_label_size = 4,
  outcome_label_vjust = -0.35,
  outcome_y_breaks = 0:5, outcome_y_limit = 5.7,
  outcome_title_colour = "#18232D", outcome_title_size = 12.5,
  outcome_axis_text_colour = "#303841", outcome_axis_text_size = 11,

  # Figure size and background
  widget_width = "100%", pair_height_px = 340,
  background_colour = "#FFFFFF", graph_padding = 0.14,
  cluster_margin = 12, cluster_border_width = 1.3,

  # Node colours
  donor_fill = "deepskyblue3", donor_border = "#00688B",
  donor_font_colour = "#FFFFFF",
  negative_fill = "#FFFFFF", negative_border = "#111111",
  negative_font_colour = "#111111",
  precursor_fill = "thistle1", precursor_border = "#BA55D3",
  precursor_font_colour = "#111111",
  alternative_fill = "snow2", alternative_border = "#73737D",
  alternative_font_colour = "#111111",
  recipient_fill = "slateblue3", recipient_border = "#473C8B",
  recipient_font_colour = "#FFFFFF",
  target_fill = "rosybrown1", target_border = "#CD5C8A",
  target_font_colour = "#111111",
  case_fill = "#FFFFFF", case_border = "#111111",
  case_font_colour = "#111111",

  # Node shape, size and text
  node_shape = "box", precursor_shape = "ellipse",
  target_shape = "doublecircle",
  node_width = 1.15, node_height = 0.40, node_padding = 0.07,
  node_border_width = 1.1, node_font_size = 12,
  label_wrap_width = 23, case_title_size = 11,
  case_title_wrap_width = 48,

  # Arrow colours and line styles
  retained_edge_colour = "#104E8B",
  confirmed_absent_edge_colour = "#111111",
  partial_edge_colour = "#BA55D3",
  not_recovered_edge_colour = "#FF3E96",
  context_edge_colour = "#73737D",
  retained_edge_width = 2.5, absent_edge_width = 2.0,
  partial_edge_width = 2.2, missing_edge_width = 2.0,
  context_edge_width = 1.5,
  arrow_head = "vee", arrow_size = 0.72,
  edge_label_size = 10, edge_label_wrap_width = 18,

  # Network spacing and font
  rank_separation = 0.90, node_separation = 0.18,
  font_name = "Arial"
)
# =======================================================================
```

```{r validation-outcome-plot, fig.width=validation_style$outcome_figure_width, fig.height=validation_style$outcome_figure_height, out.width=validation_style$outcome_output_width}
p_validation_outcomes <- ggplot(
  validation_outcomes, aes(outcome, cases, fill = outcome)
) +
  geom_col(width = validation_style$outcome_bar_width) +
  geom_text(
    aes(label = cases),
    vjust = validation_style$outcome_label_vjust,
    fontface = "bold",
    size = validation_style$outcome_label_size,
    colour = validation_style$outcome_label_colour
  ) +
  scale_fill_manual(values = validation_style$outcome_colours, drop = FALSE) +
  scale_y_continuous(
    breaks = validation_style$outcome_y_breaks,
    limits = c(0, validation_style$outcome_y_limit),
    expand = expansion(mult = c(0, 0))
  ) +
  labs(
    title = "Eight literature-defined workflow validation cases",
    subtitle = "Five formal passes; one partial pathway match; two informative non-passes",
    x = NULL, y = "Cases"
  ) +
  theme_science() +
  theme(
    legend.position = "none",
    plot.title = element_text(
      colour = validation_style$outcome_title_colour,
      size = validation_style$outcome_title_size
    ),
    axis.text = element_text(
      colour = validation_style$outcome_axis_text_colour,
      size = validation_style$outcome_axis_text_size
    )
  )

p_validation_outcomes
```

Cases 01 and 02 were matched triad controls expected to retain a
literature-supported positive branch while excluding a biologically
unsupported donor or recipient branch; the workflow reproduced both outcomes.
Cases 03--05 were independent positive controls representing
succinate-to-propionate, acetate-to-butyrate, and
lactate-to-hydrogen-sulfide cross-feeding, and all three expected relationships
were recovered. Case 06 was expected to recover the literature-described
*S. gordonii*--ornithine--*F. nucleatum* pathway, but the ornithine donor edge
was not retained. Instead, the GEMs supported a potential alternative
*S. gordonii*--L-arginine--*F. nucleatum* route to putrescine, resulting in a
partial match. Cases 07 and 08 recovered recipient-side
indole-3-propionate and deoxycholate production potential, respectively, but
did not recover the literature-described indole-3-lactate or cholate donor
edges and were classified as informative non-passes. Overall, five cases
passed formally, one produced a partial pathway match, and two exposed
donor-side limitations in the available GEMs or workflow.

**Figure V1. Pipeline-validation outcomes.** Bars give the number of the eight
literature-defined cases classified as a formal pass, partial pathway match, or
informative non-pass. Cases 01 and 02 are counted once each even though each
contains both a positive and negative branch. Colours indicate outcome class
only and do not quantify biological effect size.

The pathway panels use four edge meanings: deep-sky-blue solid arrows are
expected relationships retained by the workflow; black dashed T-bars are
expected null relationships correctly excluded; medium-orchid arrows are
recovered partial or recipient-side results; and violet-red dashed T-bars are
literature-expected edges that were not recovered. Grey dashed arrows provide
literature-route context only.

#### Validation cases 01--02

```{r validation-cases-01-02}
network_with_downloads(
  make_validation_composite(validation_case_specs[1:2], validation_style),
  "figure_V2_validation_cases_01_02"
)
```

**Figure V2. Matched triad validation controls.** Case 01 holds the recipient
constant and contrasts a retained *Bifidobacterium adolescentis* L2-32
L-lactate donor with an excluded *Paraprevotella xylaniphila* YIT 11841 donor.
Case 02 holds the donor constant and contrasts retained L-lactate use by
*Eubacterium hallii* L2-7 with the excluded negative recipient
*Phascolarctobacterium succinatutens* YIT 12067. A dashed T-bar denotes the
prespecified absent relationship and therefore contributes to a formal pass.

#### Validation cases 03--04

```{r validation-cases-03-04}
network_with_downloads(
  make_validation_composite(validation_case_specs[3:4], validation_style),
  "figure_V3_validation_cases_03_04"
)
```

**Figure V3. Independent short-chain-fatty-acid validation anchors.** Case 03
shows the retained *P. xylaniphila* YIT 11841 → succinate → *P. succinatutens*
YIT 12067 → propionate relationship. Case 04 shows the retained
*Bifidobacterium longum* NCC2705 → acetate → *Eubacterium rectale* ATCC 33656
→ butyrate relationship. Solid arrows denote retained expected components.

#### Validation case 05

```{r validation-case-05}
network_with_downloads(
  make_validation_composite(validation_case_specs[5], validation_style),
  "figure_V4_validation_case_05"
)
```

**Figure V4. Hydrogen-sulfide validation anchor.** Case 05
retained the expected *B. adolescentis* L2-32 → L-lactate → *Desulfovibrio
piger* ATCC 29098 → hydrogen-sulfide relationship.

#### Validation cases 06--08

```{r validation-cases-06-08}
network_with_downloads(
  make_validation_composite(validation_case_specs[6:8], validation_style),
  "figure_V5_validation_cases_06_08"
)
```

**Figure V5. Partial and informative non-pass validation cases.** Case 06
recovered putrescine-production potential and a GEM-supported potential
alternative *S. gordonii* CH1 → L-arginine → *F. nucleatum* ATCC 25586 route,
but did not recover the literature-described *S. gordonii* CH1 → ornithine
donor edge. Case 07 recovered indole-3-propionate-production potential in
*Clostridium sporogenes* ATCC 15579, but the donor GEM did not represent the
expected indole-3-lactate metabolite. Case 08 recovered
deoxycholate-production potential and identified cholate on the recipient
side, but cholate production was not retained for *Bacteroides
thetaiotaomicron* VPI-5482. Violet-red dashed T-bars mark the expected donor
components that were not recovered.

```{r validation-evidence-table, results='asis'}
validation_table <- validation_summary |>
  transmute(
    Case = case,
    Donor = donor,
    Precursor = precursor,
    Recipient = recipient,
    Target = target,
    `Literature expectation` = literature_expectation,
    `Workflow result` = workflow_result,
    Conclusion = conclusion,
    Reference = primary_reference
  )

validation_table$Reference <- gsub(
  "(https://doi.org/[^; ]+)",
  "<a href='\\1' target='_blank'>\\1</a>",
  validation_table$Reference,
  perl = TRUE
)

cat(as.character(tags$details(
  style = "max-width:100%;overflow:hidden;margin-bottom:1em;",
  tags$summary("Open the complete 10-row validation evidence table"),
  tags$div(
    style = "max-width:100%;overflow-x:auto;",
    HTML(kable(
      validation_table, format = "html", escape = FALSE, row.names = FALSE,
      table.attr = paste0(
        "class='table table-striped table-bordered table-condensed' ",
        "style='font-size:10.5px;width:max-content;min-width:100%;'"
      )
    ))
  )
)))
```

### CANNEX cross-feeding analysis

Sequence and model refinement yielded a final 29-GEM Core-6 panel representing
four analyzed ASVs and a 37-GEM Core-15 panel representing seven ASVs. The core
workflow evaluated 6,517 eligible different-species model tests in Core-6 and
20,184 in Core-15, retaining 2,793 (42.9%) and 6,688 (33.1%) model-level
potential edges, respectively. After species-level collapse, these corresponded
to 95 Core-6 and 554 Core-15 directional donor–precursor–recipient–target
relationships. Because the two taxon sets overlap, their records were analyzed
and reported separately rather than summed as independent relationships.

#### Taxon refinement

```{r taxon-refinement-parameters, include=FALSE}
# ================= USER-EDITABLE: FIGURE A — TAXA =====================
taxon_plot_style <- list(
  # Figure size
  figure_width = 12, figure_height = 4.8, output_width = "100%",

  # Bars and count labels
  bar_fill = manuscript_palette$deepskyblue,
  bar_border = manuscript_palette$black,
  bar_width = 0.67, bar_border_width = 0.25,
  count_label_colour = manuscript_palette$black,
  count_label_size = 3.5, count_label_hjust = -0.28,

  # Axes, facets and title
  y_upper_expansion = 0.16, y_break_count = 4,
  title_colour = "#18232D", title_size = 12.5,
  axis_text_colour = "#303841", axis_text_size = 11,
  axis_title_colour = "#303841", axis_title_size = 11,
  facet_text_colour = "#25313C", facet_text_size = 11,
  facet_fill = "#EFF2F4"
)
# ======================================================================
```

```{r taxon-refinement-plot, fig.width=taxon_plot_style$figure_width, fig.height=taxon_plot_style$figure_height, out.width=taxon_plot_style$output_width}
taxa <- data.frame(
  analysis = c(rep("Core-6", 2), rep("Core-15", 3)),
  stage = c(
    "Locked taxa", "Analyzed ASVs",
    "Locked taxa", "Genus/species eligible", ">=99% 16S match"
  ),
  count = c(6, 4, 15, 13, 7)
)
taxa$analysis <- factor(taxa$analysis, levels = c("Core-6", "Core-15"))
taxa$stage <- factor(
  taxa$stage,
  levels = rev(c("Locked taxa", "Genus/species eligible", "Analyzed ASVs", ">=99% 16S match"))
)

p_taxa <- ggplot(taxa, aes(stage, count)) +
  geom_col(
    width = taxon_plot_style$bar_width,
    fill = taxon_plot_style$bar_fill,
    colour = taxon_plot_style$bar_border,
    linewidth = taxon_plot_style$bar_border_width
  ) +
  geom_text(
    aes(label = count),
    hjust = taxon_plot_style$count_label_hjust,
    fontface = "bold",
    size = taxon_plot_style$count_label_size,
    colour = taxon_plot_style$count_label_colour
  ) +
  facet_wrap(~analysis, nrow = 1, scales = "free_x") +
  coord_flip(clip = "off") +
  scale_y_continuous(
    expand = expansion(mult = c(0, taxon_plot_style$y_upper_expansion)),
    breaks = pretty_breaks(taxon_plot_style$y_break_count)
  ) +
  labs(title = "A  Taxon refinement", x = NULL, y = "Count") +
  theme_science() +
  theme(
    legend.position = "none",
    plot.title = element_text(
      colour = taxon_plot_style$title_colour,
      size = taxon_plot_style$title_size
    ),
    axis.text = element_text(
      colour = taxon_plot_style$axis_text_colour,
      size = taxon_plot_style$axis_text_size
    ),
    axis.title = element_text(
      colour = taxon_plot_style$axis_title_colour,
      size = taxon_plot_style$axis_title_size
    ),
    strip.text = element_text(
      colour = taxon_plot_style$facet_text_colour,
      size = taxon_plot_style$facet_text_size,
      face = "bold"
    ),
    strip.background = element_rect(
      fill = taxon_plot_style$facet_fill, colour = NA
    )
  )

p_taxa
```

Core-6 retained four of the six locked ASVs for analysis. For Core-15, 13 of 15
locked ASVs had genus- or species-level assignments eligible for refinement,
and seven met the prespecified complete-coverage and ≥99% 16S identity rule.

**Figure A. Taxon refinement for the two CANNEX analysis sets.** Bars show the
number of taxa or ASVs retained at each prespecified refinement stage. Core-6
and Core-15 are shown in separate facets because they are overlapping analysis
sets and passed through different intermediate selection stages. Bar height
represents counts; it does not represent abundance or effect size.

#### GEM selection and QC

```{r gem-selection-parameters, include=FALSE}
# ================== USER-EDITABLE: FIGURE B — GEMS ====================
gem_plot_style <- list(
  # Figure size
  figure_width = 12, figure_height = 4.8, output_width = "100%",

  # Bars and count labels
  bar_fill = manuscript_palette$lightblue,
  bar_border = manuscript_palette$black,
  bar_width = 0.67, bar_border_width = 0.25,
  count_label_colour = manuscript_palette$black,
  count_label_size = 3.5, count_label_hjust = -0.28,

  # Axes, facets and title
  y_upper_expansion = 0.16, y_break_count = 4,
  title_colour = "#18232D", title_size = 12.5,
  axis_text_colour = "#303841", axis_text_size = 11,
  axis_title_colour = "#303841", axis_title_size = 11,
  facet_text_colour = "#25313C", facet_text_size = 11,
  facet_fill = "#EFF2F4"
)
# ======================================================================
```

```{r gem-selection-plot, fig.width=gem_plot_style$figure_width, fig.height=gem_plot_style$figure_height, out.width=gem_plot_style$output_width}

gems <- data.frame(
  analysis = c(rep("Core-6", 3), rep("Core-15", 2)),
  stage = c(
    "Candidate/selected GEMs", "Raw-XML QC pass", "Final GEMs",
    "Candidate/selected GEMs", "Raw-XML QC pass"
  ),
  count = c(40, 30, 29, 50, 37)
)
gems$analysis <- factor(gems$analysis, levels = c("Core-6", "Core-15"))
gems$stage <- factor(
  gems$stage,
  levels = rev(c("Candidate/selected GEMs", "Raw-XML QC pass", "Final GEMs"))
)

p_gems <- ggplot(gems, aes(stage, count)) +
  geom_col(
    width = gem_plot_style$bar_width,
    fill = gem_plot_style$bar_fill,
    colour = gem_plot_style$bar_border,
    linewidth = gem_plot_style$bar_border_width
  ) +
  geom_text(
    aes(label = count),
    hjust = gem_plot_style$count_label_hjust,
    fontface = "bold",
    size = gem_plot_style$count_label_size,
    colour = gem_plot_style$count_label_colour
  ) +
  facet_wrap(~analysis, nrow = 1, scales = "free_x") +
  coord_flip(clip = "off") +
  scale_y_continuous(
    expand = expansion(mult = c(0, gem_plot_style$y_upper_expansion)),
    breaks = pretty_breaks(gem_plot_style$y_break_count)
  ) +
  labs(title = "B  GEM selection and QC", x = NULL, y = "Count") +
  theme_science() +
  theme(
    legend.position = "none",
    plot.title = element_text(
      colour = gem_plot_style$title_colour,
      size = gem_plot_style$title_size
    ),
    axis.text = element_text(
      colour = gem_plot_style$axis_text_colour,
      size = gem_plot_style$axis_text_size
    ),
    axis.title = element_text(
      colour = gem_plot_style$axis_title_colour,
      size = gem_plot_style$axis_title_size
    ),
    strip.text = element_text(
      colour = gem_plot_style$facet_text_colour,
      size = gem_plot_style$facet_text_size,
      face = "bold"
    ),
    strip.background = element_rect(
      fill = gem_plot_style$facet_fill, colour = NA
    )
  )

p_gems
```

Of 40 candidate Core-6 GEMs, 30 passed raw XML parsing and 29 remained after
the completed sequence-refinement workflow. Of 50 sequence-selected Core-15
GEMs, 37 passed identical raw XML QC and comprised the final Core-15 panel.
Excluded models were not repaired or rewritten.

**Figure B. GEM selection and source-file quality control.** Bars report
candidate or sequence-selected AGORA2 v2.01 GEMs, models passing unmodified raw
SBML/XML parsing, and the final analysis panel where an additional refinement
stage applied. The Core-15 raw-XML-QC count is also its final panel size. The
same file-integrity rule was used for both analysis sets.

#### Core-screen retention

```{r core-screen-retention-parameters, include=FALSE}
# =============== USER-EDITABLE: FIGURE C — RETENTION ==================
retention_plot_style <- list(
  # Figure size
  figure_width = 5.5, figure_height = 2.25, output_width = "50%",

  # Stacked bars
  bar_colours = c(
    "Not retained" = manuscript_palette$lightblue,
    "Retained" = manuscript_palette$mediumorchid
  ),
  bar_border = manuscript_palette$black,
  bar_width = 0.40, bar_border_width = 0.25,

  # Labels and text
  retained_label_colour = manuscript_palette$white,
  not_retained_label_colour = manuscript_palette$black,
  label_size = 2.55, label_lineheight = 0.92,
  title_colour = "#18232D", title_size = 9.0,
  subtitle_colour = "#30303A", subtitle_size = 7.0,
  axis_text_colour = "#303841", axis_text_size = 7.0,
  axis_title_colour = "#303841", axis_title_size = 7.5,
  legend_text_colour = "#303841", legend_text_size = 7.0,
  legend_key_height_cm = 0.30
)
# ======================================================================
```

```{r core-screen-retention-plot, fig.width=retention_plot_style$figure_width, fig.height=retention_plot_style$figure_height, out.width=retention_plot_style$output_width}

screen <- data.frame(
  analysis = rep(c("Core-6", "Core-15"), each = 2),
  status = rep(c("Retained", "Not retained"), 2),
  count = c(2793, 6517 - 2793, 6688, 20184 - 6688)
)
screen$total <- ave(screen$count, screen$analysis, FUN = sum)
screen$proportion <- screen$count / screen$total
screen$label <- ifelse(
  screen$status == "Retained",
  paste0(comma(screen$count), "\n", percent(screen$proportion, accuracy = 0.1)),
  paste0(comma(screen$count), "\nnot retained")
)
screen$status <- factor(screen$status, levels = c("Not retained", "Retained"))
screen$analysis <- factor(screen$analysis, levels = c("Core-15", "Core-6"))

p_screen <- ggplot(screen, aes(analysis, count, fill = status)) +
  geom_col(
    position = "fill", width = retention_plot_style$bar_width,
    colour = retention_plot_style$bar_border,
    linewidth = retention_plot_style$bar_border_width
  ) +
  geom_text(
    aes(label = label), position = position_fill(vjust = 0.5),
    size = retention_plot_style$label_size,
    lineheight = retention_plot_style$label_lineheight,
    colour = ifelse(
      screen$status == "Retained",
      retention_plot_style$retained_label_colour,
      retention_plot_style$not_retained_label_colour
    ),
    fontface = ifelse(screen$status == "Retained", "bold", "plain")
  ) +
  coord_flip() +
  scale_y_continuous(labels = percent_format(), expand = c(0, 0)) +
  scale_fill_manual(values = retention_plot_style$bar_colours) +
  labs(
    title = "C  Eligible model tests retained by the core screen",
    subtitle = "Core-6: 2,793/6,517; Core-15: 6,688/20,184",
    x = NULL,
    y = "Eligible model tests (%)"
  ) +
  theme_science() +
  theme(
    plot.title = element_text(
      colour = retention_plot_style$title_colour,
      size = retention_plot_style$title_size
    ),
    plot.subtitle = element_text(
      colour = retention_plot_style$subtitle_colour,
      size = retention_plot_style$subtitle_size
    ),
    axis.text = element_text(
      colour = retention_plot_style$axis_text_colour,
      size = retention_plot_style$axis_text_size
    ),
    axis.title = element_text(
      colour = retention_plot_style$axis_title_colour,
      size = retention_plot_style$axis_title_size
    ),
    legend.text = element_text(
      colour = retention_plot_style$legend_text_colour,
      size = retention_plot_style$legend_text_size
    ),
    legend.key.height = grid::unit(
      retention_plot_style$legend_key_height_cm, "cm"
    )
  )

p_screen
```

The workflow retained 2,793 of 6,517 eligible Core-6 model tests (42.9%) and
6,688 of 20,184 eligible Core-15 tests (33.1%). These records represent
model-level structural-potential tests before species-level collapse.

**Figure C. Retention of eligible model-level cross-feeding tests.** Stacked
bars show the proportion of different-species donor tests retained or not
retained by the complete six-stage workflow. Labels provide exact counts and
the retained percentage within each analysis set. Core-6 and Core-15 are not
pooled because their taxon and GEM panels overlap.

#### Revised literature-classification counts

```{r exact-count-table, results='asis'}
literature_totals <- data.frame(
  Analysis = c("Core-6", "Core-15"),
  `Final relationships` = c(95, 554),
  `Case 1: donor only` = c(0, 5),
  `Case 2: both supported` = c(9, 11),
  `No qualifying donor support` = c(86, 538),
  `Any qualifying donor support` = c(9, 16),
  check.names = FALSE
)

kable(
  literature_totals,
  format = "html",
  align = c("l", "r", "r", "r", "r", "r"),
  caption = "Overall relationship counts under the revised literature rule"
) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "left"
  )

exact_counts <- data.frame(
  Analysis = c(
    rep("Core-6", 4),
    rep("Core-15", 4)
  ),
  Metabolite = rep(c("GABA", "GABA", "Butyrate", "Butyrate"), 2),
  `Root class` = rep(c("Single-root", "Multi-root"), 4),
  `Final relationships` = c(5, 44, 22, 24, 28, 293, 100, 133),
  `Case 1: donor only` = c(0, 0, 0, 0, 0, 5, 0, 0),
  `Case 2: both supported` = c(2, 2, 3, 2, 0, 5, 6, 0),
  `No qualifying donor support` = c(3, 42, 19, 22, 28, 283, 94, 133),
  `Qualifying donor support (%)` = c("40.0", "4.5", "13.6", "8.3", "0.0", "3.4", "6.0", "0.0"),
  check.names = FALSE
)

kable(
  exact_counts,
  format = "html",
  align = c("l", "l", "l", "r", "r", "r", "r", "r"),
  caption = "Re-analysed relationship counts used in the Sankey plot"
) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "left"
  )
```

“No qualifying donor support” means that the targeted search did not identify
qualifying donor-to-exact-precursor evidence. This includes relationships for
which only the recipient-to-target reaction is supported. It is an evidence-gap
label, not proof that the modeled relationship is biologically impossible.

#### Metabolite, precursor-class and literature-support Sankey

```{r sankey-data, include=FALSE}

sankey <- exact_counts |>
  transmute(
    analysis_set = Analysis,
    target_metabolite = Metabolite,
    precursor_class = recode(
      `Root class`, "Single-root" = "Single", "Multi-root" = "Multi"
    ),
    Yes = `Case 1: donor only` + `Case 2: both supported`,
    No = `No qualifying donor support`
  ) |>
  pivot_longer(
    cols = c(Yes, No),
    names_to = "literature_support",
    values_to = "count"
  ) |>
  filter(count > 0)
```

Edit the following parameter chunk when you want to change the Sankey appearance. The plotting chunk immediately after it should not need to be edited.

```{r sankey-appearance-parameters, include=FALSE}
# ====================== USER-EDITABLE SANKEY PARAMETERS ======================
sankey_style <- list(
  # Flow/ribbon colours
  flow_colours = c(
    "GABA" = "#BA55D3",
    "Butyrate" = "#00BFFF"
  ),

  # Sankey node/bar colours: all strata are intentionally white
  node_bar_colours = c(
    "Core-6" = "#FFFFFF",
    "Core-15" = "#FFFFFF",
    "GABA" = "#FFFFFF",
    "Butyrate" = "#FFFFFF",
    "Single" = "#FFFFFF",
    "Multi" = "#FFFFFF",
    "Yes" = "#FFFFFF",
    "No" = "#FFFFFF"
  ),

  # Flow/ribbon appearance
  flow_alpha = 0.40,       # 0 = transparent; 1 = opaque
  flow_width = 0.3,
  flow_curve_position = 0.47,

  # Node/bar appearance
  node_bar_width = 0.4,
  node_bar_border_colour = "#111111",
  node_bar_border_width = 0.55,

  # Text appearance
  node_label_colour = "#111111",
  node_label_size = 2.57,
  node_label_lineheight = 0.88,
  axis_text_colour = "#111111",
  axis_text_size = 11,
  title_colour = "#111111",
  title_size = 12.5,
  subtitle_colour = "#30303A",
  subtitle_size = 10.5,
  caption_colour = "#30303A",
  caption_size = 9.0,

  # Overall Sankey figure dimensions used during knitting
  figure_width = 14.2,
  figure_height = 7
)

knitr::opts_chunk$set(
  fig.width = sankey_style$figure_width,
  fig.height = sankey_style$figure_height
)
# ============================================================================
```

```{r sankey-plot}
p_sankey <- ggplot(
  sankey,
  aes(
    y = count,
    axis1 = analysis_set,
    axis2 = target_metabolite,
    axis3 = precursor_class,
    axis4 = literature_support
  )
) +
  geom_alluvium(
    aes(fill = target_metabolite),
    width = sankey_style$flow_width,
    alpha = sankey_style$flow_alpha,
    knot.pos = sankey_style$flow_curve_position
  ) +
  scale_fill_manual(values = sankey_style$flow_colours, drop = FALSE, guide = "none") +
  ggnewscale::new_scale_fill() +
  geom_stratum(
    aes(fill = after_stat(stratum)),
    width = sankey_style$node_bar_width,
    colour = sankey_style$node_bar_border_colour,
    linewidth = sankey_style$node_bar_border_width
  ) +
  geom_text(
    stat = "stratum",
    aes(
      label = paste0(
        as.character(after_stat(stratum)), "\n", comma(after_stat(count)), " (",
        percent(after_stat(prop), accuracy = 0.1), ")"
      )
    ),
    size = sankey_style$node_label_size,
    lineheight = sankey_style$node_label_lineheight,
    colour = sankey_style$node_label_colour
  ) +
  scale_x_discrete(
    limits = c(
      "Analysis set", "Target metabolite", "Precursors", "Literature support"
    ),
    expand = expansion(mult = c(0.07, 0.07))
  ) +
  scale_fill_manual(values = sankey_style$node_bar_colours, drop = FALSE, guide = "none") +
  labs(
    title = "D-E  Analysis set, metabolite, precursors and literature support",
    subtitle = paste0(
      "Ribbon width represents analysis-specific species-level relationship records; ",
      "ribbons are coloured by target metabolite"
    ),
    x = NULL,
    y = "Analysis-specific relationship records",
    caption = paste(
      strwrap(
        paste0(
          "Yes combines Case 1 (donor-to-precursor support) and Case 2 (both terminal reactions supported); No means that no qualifying donor-to-exact-precursor evidence was identified. ",
          "Recipient-to-target evidence alone is classified as No. Multi indicates that at least one retained recipient path required additional non-currency extracellular roots; it does not establish their community sources. ",
          "The combined Sankey contains analysis-specific records from overlapping Core-6 and Core-15 sets and must not be interpreted as 649 unique biological relationships."
        ),
        width = 185
      ),
      collapse = "\n"
    )
  ) +
  theme_science() +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(
      face = "bold",
      colour = sankey_style$axis_text_colour,
      size = sankey_style$axis_text_size
    ),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none",
    plot.title = element_text(
      face = "bold",
      colour = sankey_style$title_colour,
      size = sankey_style$title_size
    ),
    plot.subtitle = element_text(
      colour = sankey_style$subtitle_colour,
      size = sankey_style$subtitle_size
    ),
    plot.caption = element_text(
      colour = sankey_style$caption_colour,
      size = sankey_style$caption_size,
      hjust = 0,
      margin = margin(t = 10)
    )
  )

p_sankey
```

Across the 95 Core-6 relationships, 27 were single-precursor and 68 were
multi-precursor; nine had qualifying donor evidence and also recipient-target
support, while 86 lacked qualifying donor-to-exact-precursor evidence. Across
the 554 Core-15 relationships, 128 were single-precursor and 426 were
multi-precursor; 16 had qualifying donor evidence (five Case 1 and 11 Case 2),
while 538 lacked qualifying donor evidence. The `No` category therefore
describes the outcome of the targeted literature search rather than a negative
biological conclusion.

**Figure D–E. Distribution of CANNEX cross-feeding relationships by analysis
set, target metabolite, precursor class, and literature support.** Ribbon width
represents the number of analysis-specific species-level relationships, and
ribbon colour identifies GABA (medium orchid) or butyrate (deep sky blue).
White strata contain the category label, exact count, and within-axis
percentage. `Single` indicates that at least one retained recipient path used
only the focal precursor; `Multi` indicates that the retained paths required
additional non-currency extracellular roots. Literature support is `Yes` when
the donor-to-exact-precursor reaction is supported (Case 1 or Case 2) and `No`
when no qualifying donor evidence was identified; recipient-to-target evidence
alone remains `No`. Flow opacity is 0.4. Core-6 and Core-15 overlap and must not
be summed as independent biological relationships.

### Case-2 pathway visualisation

These two metabolite-specific networks show **Case 2 only**: direct evidence supports both the
donor-to-exact-precursor reaction and the recipient-to-target reaction. The
displayed chain is **potential donor species → focal extracellular precursor →
potential recipient species → target metabolite**. Edge colour identifies
whether a pathway came from Core-6, Core-15, or both sets. Node fill colour
identifies donor, precursor, recipient, and target roles. Because only Case 2 is
drawn, both terminal reactions meet the literature rule; the middle transfer
edge remains GEM-supported structural potential.

Single-root and multi-root relationships are both retained, but the figure does
not use `S/M` labels. A single-root relationship connects the focal precursor
directly to the recipient. For a multi-root relationship, an additional box
shows the exact displayed input combination (focal precursor plus the other
required metabolite roots). A recipient GEM counts only if that same GEM has
the exact target-production path and can import every root through an
import-capable exchange and inward transport route. The displayed combination
is selected reproducibly: fewest total roots first, then greatest complete-input
recipient-GEM support, then greatest exact-path and pathway support, and finally
alphabetical order. Every observed alternative root combination remains in the
complete Case-2 table.

```{r pathway-data-preparation, include=FALSE}
if (!exists("crossfeeding_input_paths", inherits = TRUE)) {
  stop(
    "Run the 'input-folder-configuration' chunk before this chunk."
  )
}
data_root <- crossfeeding_input_paths$data_root

pathway_inputs <- c(
  core6_workbook = file.path(
    data_root, "crossfeeding_analysis_6", "batch1_pain_gutbrain_results.xlsx"
  ),
  core6_manifest = file.path(
    data_root, "crossfeeding_analysis_6", "models_qc_passed", "model_manifest.tsv"
  ),
  core6_literature = file.path(
    data_root, "crossfeeding_analysis_6", "batch1_single_root_literature_evidence.tsv"
  ),
  core15_posthoc = file.path(
    data_root, "crossfeeding_analysis_15", "batch1_species_root_posthoc.json"
  ),
  core15_workbook = file.path(
    data_root, "crossfeeding_analysis_15", "batch1_pain_gutbrain_results.xlsx"
  ),
  core15_literature = file.path(
    data_root, "crossfeeding_analysis_15", "batch1_single_root_literature_evidence.tsv"
  )
)

if (!all(file.exists(pathway_inputs))) {
  stop(
    "Missing verified pathway input(s): ",
    paste(names(pathway_inputs)[!file.exists(pathway_inputs)], collapse = ", ")
  )
}

expected_pathway_sha256 <- c(
  core6_workbook = "3296897ac8a29282adec50d35b38305845e59ab605ed9f017031f6d8f4a101e5",
  core6_manifest = "025da099f4944b0e5969a934e1bda5a6127ce128d4b2e7926cb6ce52f6a0eaac",
  core6_literature = "d3f251b286fbeb91659860011fc4473e03d409c633541058c789b173d50c373a",
  core15_posthoc = "945e5bac6ef2429779ba00f3b86a96c63b93a996494ea91d3a82e6b640a894aa",
  core15_workbook = "2dfecf2a65985ba0dd2b384f012e071bf76cdc883358aaf3a6ccc88c683e1c3a",
  core15_literature = "306f0668ff9f88a4d7c3fb7aa6240f0b8a7821c55c96f050f57ed7eb17832703"
)
observed_pathway_sha256 <- vapply(
  pathway_inputs,
  function(path) digest(file = path, algo = "sha256", serialize = FALSE),
  character(1)
)
if (isTRUE(crossfeeding_strict_input_hash) &&
    !identical(observed_pathway_sha256, expected_pathway_sha256)) {
  changed <- names(observed_pathway_sha256)[
    observed_pathway_sha256 != expected_pathway_sha256
  ]
  stop("Verified pathway input changed: ", paste(changed, collapse = ", "))
}

fmt_n_over_n <- function(numerator, denominator) {
  sprintf("%d/%d", numerator, denominator)
}

# Reconstruct the locked Core-6 species-level relationships from the verified
# workbook with the same post-hoc rules used in the Core-6 report.
core6_results <- read_excel(
  pathway_inputs[["core6_workbook"]], sheet = "Results", na = ""
) |>
  filter(decision == "Keep")
core6_pathways <- read_excel(
  pathway_inputs[["core6_workbook"]], sheet = "Candidate Pathways", na = ""
) |>
  mutate(
    root_ids = coalesce(precursor_candidate_ids, ""),
    root_count = if_else(
      root_ids == "", 0L, str_count(root_ids, fixed(";")) + 1L
    )
  )
core6_manifest <- read.delim(
  pathway_inputs[["core6_manifest"]], check.names = FALSE,
  stringsAsFactors = FALSE
)
core6_denominators <- core6_manifest |>
  count(species, name = "species_model_n")

core6_edge_paths <- core6_results |>
  separate_rows(recipient_path_ids, sep = ";\\s*") |>
  rename(path_id = recipient_path_ids) |>
  left_join(
    select(core6_pathways, path_id, root_ids, root_count), by = "path_id"
  ) |>
  rowwise() |>
  mutate(
    focal_in_path = precursor_id %in% str_split(root_ids, ";\\s*")[[1]],
    single_root_path = focal_in_path && root_count == 1L
  ) |>
  ungroup()

if (any(is.na(core6_edge_paths$root_count)) ||
    any(!core6_edge_paths$focal_in_path)) {
  stop("A retained Core-6 edge lacks a linked path containing its focal precursor.")
}

core6_classes <- core6_edge_paths |>
  group_by(
    final_target_id, final_target, donor_species, precursor_id,
    precursor_name, recipient_species
  ) |>
  summarise(
    root_class = if_else(any(single_root_path), "single_root", "multi_root"),
    .groups = "drop"
  )

core6_relationships <- core6_edge_paths |>
  inner_join(
    core6_classes,
    by = c(
      "final_target_id", "final_target", "donor_species", "precursor_id",
      "precursor_name", "recipient_species"
    )
  ) |>
  filter(root_class == "multi_root" | single_root_path) |>
  group_by(
    final_target_id, final_target, donor_species, precursor_id,
    precursor_name, recipient_species, root_class
  ) |>
  summarise(
    donor_support_n = n_distinct(donor_model_id),
    recipient_support_n = n_distinct(recipient_model_id),
    pair_support_n = n_distinct(paste(donor_model_id, recipient_model_id, sep = "|")),
    supporting_donor_models = list(sort(unique(donor_model_id))),
    supporting_recipient_models = list(sort(unique(recipient_model_id))),
    supporting_pathway_count = n_distinct(path_id),
    .groups = "drop"
  ) |>
  left_join(core6_denominators, by = c("donor_species" = "species")) |>
  rename(donor_denominator = species_model_n) |>
  left_join(core6_denominators, by = c("recipient_species" = "species")) |>
  rename(recipient_denominator = species_model_n) |>
  mutate(
    pair_denominator = donor_denominator * recipient_denominator,
    analysis = "Core-6"
  )

# Read the already-verified Core-15 species/root post-hoc result.
core15_json <- fromJSON(
  pathway_inputs[["core15_posthoc"]], simplifyDataFrame = TRUE
)
core15_pathways <- read_excel(
  pathway_inputs[["core15_workbook"]], sheet = "Candidate Pathways", na = ""
) |>
  filter(decision == "Keep")
core15_raw <- core15_json$relationships
core15_relationships <- tibble(
  final_target_id = core15_raw$final_target_id,
  final_target = core15_raw$final_target,
  donor_species = core15_raw$donor_species,
  precursor_id = core15_raw$precursor_id,
  precursor_name = core15_raw$precursor_name,
  recipient_species = core15_raw$recipient_species,
  root_class = core15_raw$root_class,
  donor_support_n = core15_raw$donor_support$numerator,
  donor_denominator = core15_raw$donor_support$denominator,
  recipient_support_n = core15_raw$recipient_support$numerator,
  recipient_denominator = core15_raw$recipient_support$denominator,
  pair_support_n = core15_raw$model_pair_support$numerator,
  pair_denominator = core15_raw$model_pair_support$denominator,
  supporting_donor_models = core15_raw$supporting_donor_models,
  supporting_recipient_models = core15_raw$supporting_recipient_models,
  supporting_pathway_count = core15_raw$supporting_pathway_count,
  analysis = "Core-15"
)

# Donor evidence is read from the locked evidence tables. It is applied to the
# exact donor-species + precursor pair, independent of the downstream target.
core6_literature_rows <- read.delim(
  pathway_inputs[["core6_literature"]], check.names = FALSE,
  stringsAsFactors = FALSE, na.strings = ""
)
core15_literature_rows <- read.delim(
  pathway_inputs[["core15_literature"]], check.names = FALSE,
  stringsAsFactors = FALSE, na.strings = ""
)
donor_literature_pairs <- bind_rows(
  core6_literature_rows,
  core15_literature_rows
) |>
  filter(literature_class == "direct_supported") |>
  group_by(donor_species, precursor_id) |>
  summarise(
    donor_literature_citation = paste(sort(unique(citation)), collapse = "; "),
    donor_literature_doi = paste(sort(unique(doi)), collapse = "; "),
    .groups = "drop"
  )

# Recipient evidence is the accepted exact-strain or same-species evidence used
# in the combined constituent-reaction literature review.
recipient_literature_pairs <- tribble(
  ~recipient_species, ~final_target_id, ~recipient_literature_citation, ~recipient_literature_doi,
  "Agathobacter rectalis", "4abut", "Strandwitz et al. 2019", "10.1038/s41564-018-0307-3",
  "Eubacterium rectale", "4abut", "Strandwitz et al. 2019", "10.1038/s41564-018-0307-3",
  "Bacteroides ovatus", "4abut", "Otaru et al. 2021", "10.3389/fmicb.2021.656895",
  "Agathobacter rectalis", "but", "Rivière et al. 2015", "10.1128/AEM.02089-15",
  "Eubacterium rectale", "but", "Rivière et al. 2015", "10.1128/AEM.02089-15",
  "Gemmiger formicilis", "but", "Gossling and Moore 1975", "10.1099/00207713-25-2-202",
  "Roseburia hominis", "but", "Bhattacharya et al. 2022", "10.3390/microorganisms10122496"
)

all_pathway_relationships <- bind_rows(
  core6_relationships,
  core15_relationships
) |>
  left_join(
    donor_literature_pairs, by = c("donor_species", "precursor_id")
  ) |>
  left_join(
    recipient_literature_pairs,
    by = c("recipient_species", "final_target_id")
  ) |>
  mutate(
    donor_literature_supported = !is.na(donor_literature_citation),
    recipient_literature_supported = !is.na(recipient_literature_citation),
    literature_case = case_when(
      donor_literature_supported & recipient_literature_supported ~
        "Case 2: both supported",
      donor_literature_supported ~ "Case 1: donor only",
      TRUE ~ "No qualifying donor support"
    ),
    donor_support = fmt_n_over_n(donor_support_n, donor_denominator),
    recipient_support = fmt_n_over_n(
      recipient_support_n, recipient_denominator
    ),
    pair_support = fmt_n_over_n(pair_support_n, pair_denominator)
  ) |>
  arrange(analysis, final_target, root_class, donor_species,
          precursor_name, recipient_species)

case2_pathway_relationships <- all_pathway_relationships |>
  filter(literature_case == "Case 2: both supported")
case1_pathway_relationships <- all_pathway_relationships |>
  filter(literature_case == "Case 1: donor only")
no_donor_support_pathway_relationships <- all_pathway_relationships |>
  filter(literature_case == "No qualifying donor support")

# Recover the complete extracellular root sets from the locked candidate-pathway
# sheets and join them to the Step-5 recipient import audits. A recipient model
# supports a complete input set only when the same GEM contains the exact path
# and every root has exchange import plus inward transport. For each Case-2
# relationship, select one eligible display set deterministically: minimum root
# count, maximum complete-input recipient-model support, maximum exact-path
# support, maximum pathway support, then alphabetical order.
candidate_root_paths <- bind_rows(
  core6_pathways |>
    transmute(
      analysis = "Core-6", final_target_id = target_id,
      recipient_model_id = model_id, path_id,
      root_set_ids = coalesce(precursor_candidate_ids, ""),
      root_set_names = coalesce(precursor_candidate_names, "")
    ),
  core15_pathways |>
    transmute(
      analysis = "Core-15", final_target_id = target_id,
      recipient_model_id = model_id, path_id,
      root_set_ids = coalesce(precursor_candidate_ids, ""),
      root_set_names = coalesce(precursor_candidate_names, "")
    )
) |>
  mutate(
    root_count = if_else(
      root_set_ids == "", 0L,
      str_count(root_set_ids, fixed(";")) + 1L
    ),
    root_set_label = str_replace_all(root_set_names, ";\\s*", " + ")
  )

consumer_import_audits <- bind_rows(
  read_excel(
    pathway_inputs[["core6_workbook"]], sheet = "Consumer Audits", na = ""
  ) |>
    transmute(
      analysis = "Core-6", final_target_id = target_id,
      recipient_model_id = model_id,
      importable_precursor_ids = coalesce(
        as.character(importable_precursor_ids), ""
      )
    ),
  read_excel(
    pathway_inputs[["core15_workbook"]], sheet = "Consumer Audits", na = ""
  ) |>
    transmute(
      analysis = "Core-15", final_target_id = target_id,
      recipient_model_id = model_id,
      importable_precursor_ids = coalesce(
        as.character(importable_precursor_ids), ""
      )
    )
)

case2_root_options <- case2_pathway_relationships |>
  select(
    analysis, final_target_id, donor_species, precursor_id,
    recipient_species, root_class, recipient_denominator,
    supporting_recipient_models
  ) |>
  unnest_longer(
    supporting_recipient_models, values_to = "recipient_model_id"
  ) |>
  inner_join(
    candidate_root_paths,
    by = c("analysis", "final_target_id", "recipient_model_id")
  ) |>
  inner_join(
    consumer_import_audits,
    by = c("analysis", "final_target_id", "recipient_model_id")
  ) |>
  rowwise() |>
  mutate(
    focal_in_root_set = precursor_id %in%
      str_split(root_set_ids, ";\\s*")[[1]],
    all_roots_importable = all(
      str_split(root_set_ids, ";\\s*")[[1]] %in%
        str_split(importable_precursor_ids, ";\\s*")[[1]]
    )
  ) |>
  ungroup() |>
  filter(
    focal_in_root_set,
    (root_class == "single_root" & root_count == 1L) |
      (root_class == "multi_root" & root_count > 1L)
  ) |>
  group_by(
    analysis, final_target_id, donor_species, precursor_id,
    recipient_species, root_class, recipient_denominator,
    root_set_ids, root_set_names, root_set_label, root_count
  ) |>
  summarise(
    root_set_model_n = n_distinct(recipient_model_id),
    root_set_path_n = n_distinct(path_id),
    root_set_models = list(sort(unique(recipient_model_id))),
    root_set_importable_model_n = n_distinct(
      recipient_model_id[all_roots_importable]
    ),
    root_set_importable_models = list(sort(unique(
      recipient_model_id[all_roots_importable]
    ))),
    .groups = "drop"
  )

case2_root_summaries <- case2_root_options |>
  arrange(
    analysis, final_target_id, donor_species, precursor_id,
    recipient_species, root_count, desc(root_set_importable_model_n),
    desc(root_set_model_n), desc(root_set_path_n), root_set_label
  ) |>
  group_by(
    analysis, final_target_id, donor_species, precursor_id,
    recipient_species
  ) |>
  summarise(
    all_observed_root_sets = paste0(
      root_set_label, " [exact path: ", root_set_model_n, "/",
      recipient_denominator, "; every root importable: ",
      root_set_importable_model_n, "/", root_set_model_n,
      "]", collapse = " | "
    ),
    observed_root_set_n = n(),
    import_eligible_root_set_n = sum(root_set_importable_model_n > 0L),
    .groups = "drop"
  )

case2_display_roots <- case2_root_options |>
  filter(root_set_importable_model_n > 0L) |>
  arrange(
    analysis, final_target_id, donor_species, precursor_id,
    recipient_species, root_count, desc(root_set_importable_model_n),
    desc(root_set_model_n), desc(root_set_path_n), root_set_label
  ) |>
  group_by(
    analysis, final_target_id, donor_species, precursor_id,
    recipient_species
  ) |>
  mutate(root_set_rank = row_number()) |>
  ungroup() |>
  filter(root_set_rank == 1L) |>
  select(
    analysis, final_target_id, donor_species, precursor_id,
    recipient_species, displayed_root_set_ids = root_set_ids,
    displayed_root_set = root_set_label,
    displayed_root_count = root_count,
    displayed_root_path_support_n = root_set_model_n,
    displayed_root_support_n = root_set_importable_model_n,
    displayed_root_path_n = root_set_path_n,
    displayed_root_models = root_set_importable_models
  )

case2_pathway_relationships <- case2_pathway_relationships |>
  left_join(
    case2_display_roots,
    by = c(
      "analysis", "final_target_id", "donor_species", "precursor_id",
      "recipient_species"
    )
  ) |>
  left_join(
    case2_root_summaries,
    by = c(
      "analysis", "final_target_id", "donor_species", "precursor_id",
      "recipient_species"
    )
  )

if (any(is.na(case2_pathway_relationships$displayed_root_set))) {
  stop(
    "A Case-2 relationship lacks an exact root set for which every root is importable."
  )
}

displayed_multiroot_import_check <- case2_pathway_relationships |>
  filter(displayed_root_count > 1L) |>
  distinct(
    analysis, final_target_id, recipient_species, displayed_root_set,
    displayed_root_path_support_n, displayed_root_support_n,
    recipient_denominator
  ) |>
  arrange(analysis, final_target_id, recipient_species, displayed_root_set)
expected_multiroot_import_check <- tribble(
  ~analysis, ~final_target_id, ~recipient_species, ~displayed_root_set,
  ~displayed_root_path_support_n, ~displayed_root_support_n,
  ~recipient_denominator,
  "Core-6", "4abut", "Agathobacter rectalis", "Fumarate + Ammonium",
  8L, 8L, 10L,
  "Core-6", "but", "Agathobacter rectalis", "Acetate + L-glutamate",
  3L, 3L, 10L,
  "Core-15", "4abut", "Bacteroides ovatus", "Ammonium + Succinate",
  2L, 2L, 4L,
  "Core-15", "4abut", "Eubacterium rectale", "Ammonium + Succinate",
  7L, 7L, 10L
) |>
  arrange(analysis, final_target_id, recipient_species, displayed_root_set)
stopifnot(identical(
  displayed_multiroot_import_check,
  expected_multiroot_import_check
))

case2_normalized_keys <- case2_pathway_relationships |>
  mutate(
    donor_species_display = recode(
      donor_species, "Eubacterium rectale" = "Agathobacter rectalis"
    ),
    recipient_species_display = recode(
      recipient_species, "Eubacterium rectale" = "Agathobacter rectalis"
    )
  ) |>
  count(
    final_target_id, donor_species_display, precursor_id,
    recipient_species_display,
    name = "analysis_records"
  )
case2_unique_by_target <- case2_normalized_keys |>
  count(final_target_id, name = "unique_pathways")

# Fail loudly if the relationship-level reconstruction no longer matches the
# locked post-hoc results and the accepted literature rule.
pathway_count_check <- all_pathway_relationships |>
  count(analysis, final_target, root_class, literature_case, name = "n")
expected_pathway_count_check <- tribble(
  ~analysis, ~final_target, ~root_class, ~literature_case, ~n,
  "Core-6", "GABA", "single_root", "Case 2: both supported", 2L,
  "Core-6", "GABA", "single_root", "No qualifying donor support", 3L,
  "Core-6", "GABA", "multi_root", "Case 2: both supported", 2L,
  "Core-6", "GABA", "multi_root", "No qualifying donor support", 42L,
  "Core-6", "Butyrate", "single_root", "Case 2: both supported", 3L,
  "Core-6", "Butyrate", "single_root", "No qualifying donor support", 19L,
  "Core-6", "Butyrate", "multi_root", "Case 2: both supported", 2L,
  "Core-6", "Butyrate", "multi_root", "No qualifying donor support", 22L,
  "Core-15", "GABA", "single_root", "No qualifying donor support", 28L,
  "Core-15", "GABA", "multi_root", "Case 1: donor only", 5L,
  "Core-15", "GABA", "multi_root", "Case 2: both supported", 5L,
  "Core-15", "GABA", "multi_root", "No qualifying donor support", 283L,
  "Core-15", "Butyrate", "single_root", "Case 2: both supported", 6L,
  "Core-15", "Butyrate", "single_root", "No qualifying donor support", 94L,
  "Core-15", "Butyrate", "multi_root", "No qualifying donor support", 133L
) |>
  arrange(analysis, final_target, root_class, literature_case)
stopifnot(
  identical(
    pathway_count_check |>
      arrange(analysis, final_target, root_class, literature_case),
    expected_pathway_count_check
  ),
  nrow(case2_pathway_relationships) == 20L,
  sum(case2_pathway_relationships$displayed_root_count == 1L) == 11L,
  sum(case2_pathway_relationships$displayed_root_count > 1L) == 9L,
  nrow(case1_pathway_relationships) == 5L,
  nrow(no_donor_support_pathway_relationships) == 624L,
  nrow(case2_normalized_keys) == 20L,
  all(case2_normalized_keys$analysis_records == 1L),
  case2_unique_by_target$unique_pathways[
    case2_unique_by_target$final_target_id == "4abut"
  ] == 9L,
  case2_unique_by_target$unique_pathways[
    case2_unique_by_target$final_target_id == "but"
  ] == 11L
)

actual_sankey_check <- all_pathway_relationships |>
  mutate(
    precursor_class = recode(
      root_class, single_root = "Single-root", multi_root = "Multi-root"
    ),
    precursor_class = recode(
      precursor_class, "Single-root" = "Single", "Multi-root" = "Multi"
    ),
    literature_support = if_else(
      donor_literature_supported, "Yes", "No"
    )
  ) |>
  count(
    analysis_set = analysis,
    target_metabolite = final_target,
    precursor_class,
    literature_support,
    name = "count"
  ) |>
  arrange(
    analysis_set, target_metabolite, precursor_class, literature_support
  )
expected_sankey_check <- sankey |>
  mutate(count = as.integer(count)) |>
  arrange(
    analysis_set, target_metabolite, precursor_class, literature_support
  )
stopifnot(identical(actual_sankey_check, as_tibble(expected_sankey_check)))

dot_escape <- function(x) {
  x <- gsub("\\\\", "\\\\\\\\", as.character(x))
  x <- gsub('"', '\\"', x, fixed = TRUE)
  gsub("\n", "\\n", x, fixed = TRUE)
}

wrap_dot_label <- function(x, width = 18L) {
  dot_escape(str_wrap(as.character(x), width = width))
}

edge_width_from_frequency <- function(frequency, style) {
  if (frequency <= style$frequency_break_low) return(style$edge_width_low)
  if (frequency <= style$frequency_break_mid) return(style$edge_width_mid)
  style$edge_width_high
}

analysis_membership <- function(x) {
  values <- sort(unique(as.character(x)))
  if ("Both" %in% values || all(c("Core-6", "Core-15") %in% values)) {
    return("Both")
  }
  values[[1]]
}

analysis_edge_colour <- function(analysis_class, style) {
  switch(
    analysis_class,
    "Core-6" = style$core6_edge_colour,
    "Core-15" = style$core15_edge_colour,
    "Both" = style$both_sets_edge_colour,
    style$both_sets_edge_colour
  )
}

make_supported_pathway_network <- function(
    relationships, target_id, target_label, style) {
  rel_raw <- relationships |>
    filter(final_target_id == target_id) |>
    mutate(
      donor_species = recode(
        donor_species, "Eubacterium rectale" = "Agathobacter rectalis"
      ),
      recipient_species = recode(
        recipient_species, "Eubacterium rectale" = "Agathobacter rectalis"
      )
    )
  rel <- rel_raw |>
    group_by(
      final_target_id, final_target, donor_species, precursor_id,
      precursor_name, recipient_species, displayed_root_set_ids,
      displayed_root_set, displayed_root_count
    ) |>
    summarise(
      analysis = analysis_membership(analysis),
      donor_denominator = max(donor_denominator),
      recipient_denominator = max(recipient_denominator),
      supporting_donor_models = list(sort(unique(unlist(
        supporting_donor_models, use.names = FALSE
      )))),
      supporting_recipient_models = list(sort(unique(unlist(
        supporting_recipient_models, use.names = FALSE
      )))),
      displayed_root_models = list(sort(unique(unlist(
        displayed_root_models, use.names = FALSE
      )))),
      donor_literature_supported = any(donor_literature_supported),
      recipient_literature_supported = any(recipient_literature_supported),
      .groups = "drop"
    )
  if (!nrow(rel)) stop("No Case-2 relationships for this panel.")

  donors <- sort(unique(rel$donor_species))
  precursors <- rel |>
    distinct(precursor_id, precursor_name) |>
    arrange(precursor_name)
  recipients <- sort(unique(rel$recipient_species))

  donor_ids <- setNames(paste0("d", seq_along(donors)), donors)
  precursor_ids <- setNames(
    paste0("p", seq_len(nrow(precursors))), precursors$precursor_id
  )
  recipient_ids <- setNames(paste0("r", seq_along(recipients)), recipients)

  rel <- rel |>
    mutate(
      input_set_key = if_else(
        displayed_root_count > 1L,
        paste(
          precursor_id, displayed_root_set_ids, recipient_species,
          sep = "|"
        ),
        NA_character_
      )
    )
  input_sets <- rel |>
    filter(displayed_root_count > 1L) |>
    group_by(input_set_key, displayed_root_set, recipient_species) |>
    summarise(
      analysis_class = analysis_membership(analysis),
      support_models = list(sort(unique(unlist(
        displayed_root_models, use.names = FALSE
      )))),
      denominator = max(recipient_denominator),
      .groups = "drop"
    ) |>
    arrange(displayed_root_set, recipient_species) |>
    mutate(node_id = paste0("q", row_number()))
  input_set_ids <- setNames(input_sets$node_id, input_sets$input_set_key)
  rel <- rel |>
    mutate(
      input_set_node = unname(input_set_ids[input_set_key])
    )

  donor_edges <- rel |>
    group_by(donor_species, precursor_id) |>
    summarise(
      support_models = list(sort(unique(unlist(
        supporting_donor_models, use.names = FALSE
      )))),
      denominator = first(donor_denominator),
      analysis_class = analysis_membership(analysis),
      .groups = "drop"
    ) |>
    mutate(
      numerator = lengths(support_models),
      frequency = numerator / denominator,
      from = unname(donor_ids[donor_species]),
      to = unname(precursor_ids[precursor_id]),
      label = sprintf("%d/%d", numerator, denominator),
      edge_group = "donor"
    )

  single_input_edges <- rel |>
    filter(displayed_root_count == 1L) |>
    group_by(precursor_id, recipient_species) |>
    summarise(
      support_models = list(sort(unique(unlist(
        displayed_root_models, use.names = FALSE
      )))),
      denominator = first(recipient_denominator),
      analysis_class = analysis_membership(analysis),
      .groups = "drop"
    ) |>
    mutate(
      numerator = lengths(support_models),
      frequency = numerator / denominator,
      from = unname(precursor_ids[precursor_id]),
      to = unname(recipient_ids[recipient_species]),
      label = sprintf("%d/%d", numerator, denominator),
      edge_group = "single_input"
    )

  combined_input_entry_edges <- rel |>
    filter(displayed_root_count > 1L) |>
    group_by(precursor_id, input_set_node) |>
    summarise(
      support_models = list(sort(unique(unlist(
        displayed_root_models, use.names = FALSE
      )))),
      denominator = first(recipient_denominator),
      analysis_class = analysis_membership(analysis),
      .groups = "drop"
    ) |>
    mutate(
      numerator = lengths(support_models),
      frequency = numerator / denominator,
      from = unname(precursor_ids[precursor_id]),
      to = input_set_node,
      label = "",
      edge_group = "combined_input_entry"
    )

  combined_input_exit_edges <- input_sets |>
    mutate(
      numerator = lengths(support_models),
      frequency = numerator / denominator,
      from = node_id,
      to = unname(recipient_ids[recipient_species]),
      label = sprintf("%d/%d", numerator, denominator),
      edge_group = "combined_input_exit"
    )

  recipient_edges <- rel |>
    group_by(recipient_species) |>
    summarise(
      support_models = list(sort(unique(unlist(
        supporting_recipient_models, use.names = FALSE
      )))),
      denominator = first(recipient_denominator),
      analysis_class = analysis_membership(analysis),
      .groups = "drop"
    ) |>
    mutate(
      numerator = lengths(support_models),
      frequency = numerator / denominator,
      from = unname(recipient_ids[recipient_species]),
      to = "target",
      label = sprintf("%d/%d", numerator, denominator),
      edge_group = "recipient"
    )

  edges <- bind_rows(
    donor_edges, single_input_edges, combined_input_entry_edges,
    combined_input_exit_edges, recipient_edges
  )

  donor_nodes <- vapply(
    donors,
    function(label) sprintf(
      paste0(
        '%s [label="%s",shape="%s",style="rounded,filled",',
        paste0(
          'fillcolor="%s",color="%s",fontcolor="%s",penwidth=%.2f,',
          'width=%.2f,height=%.2f];'
        )
      ),
      donor_ids[[label]], wrap_dot_label(label),
      style$donor_shape,
      style$donor_fill, style$donor_border,
      style$donor_font_colour, style$node_border_width,
      style$node_width, style$node_height
    ),
    character(1)
  )
  precursor_nodes <- vapply(
    seq_len(nrow(precursors)),
    function(i) sprintf(
      paste0(
        '%s [label="%s",shape="%s",style=filled,',
        'fillcolor="%s",color="%s",fontcolor="%s",penwidth=%.2f];'
      ),
      precursor_ids[[precursors$precursor_id[[i]]]],
      wrap_dot_label(precursors$precursor_name[[i]]),
      style$precursor_shape,
      style$precursor_fill, style$precursor_border,
      style$precursor_font_colour, style$node_border_width
    ),
    character(1)
  )
  input_set_nodes <- vapply(
    seq_len(nrow(input_sets)),
    function(i) sprintf(
      paste0(
        '%s [label="%s",shape="%s",style="rounded,filled",',
        paste0(
          'fillcolor="%s",color="%s",fontcolor="%s",penwidth=%.2f,',
          'width=%.2f,height=%.2f];'
        )
      ),
      input_sets$node_id[[i]],
      wrap_dot_label(input_sets$displayed_root_set[[i]], width = 24L),
      style$input_set_shape,
      style$input_set_fill, style$input_set_border,
      style$input_set_font_colour, style$node_border_width,
      style$input_set_width, style$input_set_height
    ),
    character(1)
  )
  recipient_nodes <- vapply(
    recipients,
    function(label) sprintf(
      paste0(
        '%s [label="%s",shape="%s",style="rounded,filled",',
        'fillcolor="%s",color="%s",fontcolor="%s",penwidth=%.2f];'
      ),
      recipient_ids[[label]], wrap_dot_label(label),
      style$recipient_shape,
      style$recipient_fill, style$recipient_border,
      style$recipient_font_colour, style$node_border_width
    ),
    character(1)
  )

  edge_lines <- vapply(
    seq_len(nrow(edges)),
    function(i) {
      frequency <- edges$frequency[[i]]
      colour <- analysis_edge_colour(edges$analysis_class[[i]], style)
      colour <- alpha(colour, style$edge_alpha)
      sprintf(
        paste0(
          '%s -> %s [label="%s",color="%s",fontcolor="%s",',
          'fontsize=%.1f,penwidth=%.2f,arrowsize=%.2f,arrowhead="%s",',
          'style="%s"];'
        ),
        edges$from[[i]], edges$to[[i]], dot_escape(edges$label[[i]]),
        colour, style$edge_label_colour, style$edge_label_size,
        edge_width_from_frequency(frequency, style),
        style$arrow_size, style$arrow_head, style$edge_line_style
      )
    },
    character(1)
  )

  title <- sprintf(
    "%s: %d unique Case-2 pathways across Core-6 and Core-15",
    target_label, nrow(rel)
  )
  dot <- paste0(
    "digraph G {",
    sprintf(
      paste0(
        'graph [rankdir=LR,bgcolor="%s",pad=%.2f,nodesep=%.2f,',
        'ranksep=%.2f,splines="%s",fontname="%s",labelloc=t,',
        'label="%s",fontcolor="%s",fontsize=%.1f];'
      ),
      style$background_colour, style$graph_padding,
      style$node_separation, style$rank_separation, style$spline_type,
      style$font_name, dot_escape(title), style$title_colour,
      style$title_size
    ),
    sprintf(
      paste0(
        'node [fontname="%s",fontsize=%.1f,margin=%.2f,',
        'width=%.2f,height=%.2f];'
      ),
      style$font_name, style$node_font_size, style$node_padding,
      style$node_width, style$node_height
    ),
    sprintf('edge [fontname="%s"];', style$font_name),
    paste(donor_nodes, collapse = "\n"),
    paste(precursor_nodes, collapse = "\n"),
    paste(input_set_nodes, collapse = "\n"),
    paste(recipient_nodes, collapse = "\n"),
    sprintf(
      paste0(
        'target [label="%s",shape="%s",style=filled,',
        'fillcolor="%s",color="%s",fontcolor="%s",penwidth=%.2f];'
      ),
      dot_escape(target_label), style$target_shape,
      style$target_fill, style$target_border,
      style$target_font_colour, style$target_border_width
    ),
    sprintf(
      paste0(
        'legend [label=<<TABLE BORDER="0" CELLBORDER="0" CELLSPACING="2">',
        '<TR><TD><FONT COLOR="%s">━━━━</FONT></TD><TD ALIGN="LEFT">Core-6</TD></TR>',
        '<TR><TD><FONT COLOR="%s">━━━━</FONT></TD><TD ALIGN="LEFT">Core-15</TD></TR>',
        '<TR><TD><FONT COLOR="%s">━━━━</FONT></TD><TD ALIGN="LEFT">Both sets</TD></TR>',
        '<TR><TD COLSPAN="2" ALIGN="LEFT">All displayed terminal reactions meet Case 2</TD></TR>',
        '<TR><TD COLSPAN="2" ALIGN="LEFT">n/N = supporting/tested GEMs</TD></TR>',
        '<TR><TD COLSPAN="2" ALIGN="LEFT">Snow-grey box = complete required input combination</TD></TR>',
        '<TR><TD COLSPAN="2" ALIGN="LEFT">Snow-grey-box n/N = exact path plus import of every root / tested GEMs</TD></TR>',
        '<TR><TD COLSPAN="2" ALIGN="LEFT">Selection: fewest inputs, then greatest eligible GEM support</TD></TR>',
        '<TR><TD COLSPAN="2" ALIGN="LEFT">All alternative combinations are retained in the table</TD></TR>',
        '</TABLE>>,shape=box,style="rounded,filled",',
        'fillcolor="%s",color="%s",fontcolor="%s",fontsize=%.1f];'
      ),
      style$core6_edge_colour, style$core15_edge_colour,
      style$both_sets_edge_colour,
      style$legend_fill, style$legend_border,
      style$legend_font_colour, style$legend_font_size
    ),
    "{rank=same;", paste(unname(donor_ids), collapse = ";"), ";legend;}",
    "{rank=same;", paste(unname(precursor_ids), collapse = ";"), ";}",
    if (nrow(input_sets)) {
      paste0("{rank=same;", paste(input_sets$node_id, collapse = ";"), ";}")
    } else {
      ""
    },
    "{rank=same;", paste(unname(recipient_ids), collapse = ";"), ";}",
    paste(edge_lines, collapse = "\n"),
    "}"
  )

  grViz(
    dot,
    width = style$widget_width,
    height = style$widget_height_px
  )
}

pathway_table_view <- function(data) {
  data |>
    transmute(
      Analysis = analysis,
      Target = final_target,
      `Potential donor species` = donor_species,
      `Focal extracellular precursor` = precursor_name,
      `Potential recipient species` = recipient_species,
      `Root class` = recode(
        root_class, single_root = "Single-root", multi_root = "Multi-root"
      ),
      `Literature class` = literature_case,
      `Donor GEM support` = donor_support,
      `Recipient GEM support` = recipient_support,
      `Model-pair support` = pair_support,
      `Supporting paths` = supporting_pathway_count,
      `Donor citation` = coalesce(donor_literature_citation, "—"),
      `Recipient citation` = coalesce(recipient_literature_citation, "—")
    )
}

case2_pathway_table_view <- function(data) {
  data |>
    transmute(
      Analysis = analysis,
      Target = final_target,
      `Potential donor species` = donor_species,
      `Focal extracellular precursor` = precursor_name,
      `Displayed recipient input set` = displayed_root_set,
      `Displayed exact-path support` = fmt_n_over_n(
        displayed_root_path_support_n, recipient_denominator
      ),
      `Every root importable among exact-path GEMs` = fmt_n_over_n(
        displayed_root_support_n, displayed_root_path_support_n
      ),
      `Eligible complete-input support` = fmt_n_over_n(
        displayed_root_support_n, recipient_denominator
      ),
      `Observed input sets` = observed_root_set_n,
      `Input sets with at least one eligible GEM` = import_eligible_root_set_n,
      `All observed input sets and support` = all_observed_root_sets,
      `Potential recipient species` = recipient_species,
      `Literature class` = literature_case,
      `Donor GEM support` = donor_support,
      `Recipient GEM support` = recipient_support,
      `Model-pair support` = pair_support,
      `Donor citation` = coalesce(donor_literature_citation, "—"),
      `Recipient citation` = coalesce(recipient_literature_citation, "—")
    )
}

make_expandable_table <- function(summary_text, data) {
  tags$details(
    style = "max-width:100%;overflow:hidden;margin-bottom:1em;",
    tags$summary(summary_text),
    tags$div(
      style = "max-width:100%;overflow-x:auto;",
      HTML(kable(
        data, format = "html", escape = TRUE, row.names = FALSE,
        table.attr = paste0(
          "class='table table-striped table-bordered table-condensed' ",
          "style='font-size:10.5px;width:max-content;min-width:100%;'"
        )
      ))
    )
  )
}
```

#### Combined GABA network

This figure contains nine unique Case-2 pathways after combining the two analysis
sets. No complete GABA pathway was duplicated between Core-6 and Core-15 after
normalizing *Eubacterium rectale* to *Agathobacter rectalis* for display.

The parameter chunk immediately below controls only the GABA network.

```{r gaba-pathway-parameters, include=FALSE}
# ===================== USER-EDITABLE: GABA PATHWAY ======================
gaba_pathway_style <- list(
  # Figure size and background
  widget_width = "100%", widget_height_px = 680,
  background_colour = "#FFFFFF", graph_padding = 0.25,

  # Node/box colours
  donor_fill = "deepskyblue3", donor_border = "#00688B",
  precursor_fill = "thistle1", precursor_border = "#BA55D3",
  input_set_fill = "snow2", input_set_border = "#73737D",
  recipient_fill = "slateblue1", recipient_border = "#473C8B",
  target_fill = "#FFC1C1", target_border = "#CD5C8A",
  donor_shape = "box", precursor_shape = "ellipse",
  input_set_shape = "box",
  recipient_shape = "box", target_shape = "doublecircle",

  # Analysis-set colours used for arrows/lines
  core6_edge_colour = "lightskyblue",
  core15_edge_colour = "steelblue4",
  both_sets_edge_colour = "slateblue1",

  # Text colours and sizes
  donor_font_colour = "#FFFFFF", precursor_font_colour = "#111111",
  recipient_font_colour = "#FFFFFF", input_set_font_colour = "#111111",
  target_font_colour = "#111111",
  node_font_size = 12.0,
  edge_label_colour = "#111111", edge_label_size = 10,
  title_colour = "#111111", title_size = 18,
  legend_font_colour = "#30303A", legend_font_size = 8.5,

  # Node/box size and border thickness
  node_width = 1.45, node_height = 0.42, node_padding = 0.09,
  input_set_width = 1.75, input_set_height = 0.50,
  node_border_width = 1.1, target_border_width = 2.1,

  # Arrow/line transparency and thickness
  edge_alpha = 0.90,
  frequency_break_low = 0.30, frequency_break_mid = 0.70,
  edge_width_low = 1.0, edge_width_mid = 2.6, edge_width_high = 5.0,
  arrow_head = "vee", arrow_size = 0.78,
  edge_line_style = "solid",

  # Network spacing and legend box
  spline_type = "spline", rank_separation = 1.20, node_separation = 0.20,
  font_name = "Arial",
  legend_fill = "#FFFFFF", legend_border = "#111111"
)
# =======================================================================
```

```{r gaba-pathway-plot}
network_with_downloads(
  make_supported_pathway_network(
    case2_pathway_relationships, "4abut", "GABA", gaba_pathway_style
  ),
  "figure_F_GABA_case2_network"
)
```

The GABA network contained four Core-6 Case-2 relationships (two
single-precursor and two multi-precursor) and five Core-15 Case-2
relationships (all multi-precursor). No complete relationship was duplicated
between the two sets after taxonomic display-name normalization. For every
displayed multi-precursor route, at least one exact-path recipient GEM could
import and transport every metabolite in the displayed input set.

**Figure F. Literature-supported Case-2 pathways ending in GABA.** The network
shows only relationships for which both donor-to-exact-precursor and
recipient-to-GABA evidence met the prespecified literature rule. Light-sky-blue
edges denote Core-6, steel-blue edges denote Core-15, and slate-blue edges
would denote a relationship present in both sets. Node colours distinguish
potential donors (deep sky blue), focal precursors (thistle), potential
recipients (slate blue), and GABA (pale violet red). A snow-grey intermediate
box lists the complete input combination for a displayed multi-precursor path.
Edge labels report supporting/tested GEMs (`n/N`); line width denotes low
(≤30%), intermediate (>30% to ≤70%), or high (>70%) strain-panel support.
These are structural-potential relationships, not direct evidence of in vivo
metabolite transfer.

#### Combined butyrate network

This figure contains eleven unique Case-2 pathways after combining the two
analysis sets. No complete butyrate pathway was duplicated between Core-6 and
Core-15 after taxonomic display-name normalization.

The parameter chunk immediately below controls only the butyrate network.

```{r butyrate-pathway-parameters, include=FALSE}
# =================== USER-EDITABLE: BUTYRATE PATHWAY ====================
butyrate_pathway_style <- list(
  # Figure size and background
  widget_width = "100%", widget_height_px = 720,
  background_colour = "#FFFFFF", graph_padding = 0.25,

  # Node/box colours
  donor_fill = "deepskyblue3", donor_border = "#00688B",
  precursor_fill = "thistle1", precursor_border = "#BA55D3",
  input_set_fill = "snow2", input_set_border = "#73737D",
  recipient_fill = "slateblue1", recipient_border = "#473C8B",
  target_fill = "#FFC1C1", target_border = "#CD5C8A",
  donor_shape = "box", precursor_shape = "ellipse",
  input_set_shape = "box",
  recipient_shape = "box", target_shape = "doublecircle",

  # Analysis-set colours used for arrows/lines
  core6_edge_colour = "lightskyblue",
  core15_edge_colour = "steelblue4",
  both_sets_edge_colour = "slateblue1",

  # Text colours and sizes
  donor_font_colour = "#FFFFFF", precursor_font_colour = "#111111",
  recipient_font_colour = "#FFFFFF", input_set_font_colour = "#111111",
  target_font_colour = "#111111",
  node_font_size = 10.0,
  edge_label_colour = "#111111", edge_label_size = 10,
  title_colour = "#111111", title_size = 18,
  legend_font_colour = "#30303A", legend_font_size = 8.5,

  # Node/box size and border thickness
  node_width = 1.45, node_height = 0.42, node_padding = 0.09,
  input_set_width = 1.75, input_set_height = 0.50,
  node_border_width = 1.1, target_border_width = 2.1,

  # Arrow/line transparency and thickness
  edge_alpha = 0.90,
  frequency_break_low = 0.30, frequency_break_mid = 0.70,
  edge_width_low = 1.0, edge_width_mid = 2.6, edge_width_high = 5.0,
  arrow_head = "vee", arrow_size = 0.78,
  edge_line_style = "solid",

  # Network spacing and legend box
  spline_type = "spline", rank_separation = 1.20, node_separation = 0.20,
  font_name = "Arial",
  legend_fill = "#FFFFFF", legend_border = "#111111"
)
# =======================================================================
```

```{r butyrate-pathway-plot}
network_with_downloads(
  make_supported_pathway_network(
    case2_pathway_relationships, "but", "Butyrate", butyrate_pathway_style
  ),
  "figure_G_butyrate_case2_network"
)
```

The butyrate network contained five Core-6 Case-2 relationships (three
single-precursor and two multi-precursor) and six Core-15 Case-2 relationships
(all single-precursor). No complete relationship was duplicated between the
two sets after taxonomic display-name normalization. Each displayed
multi-precursor input set passed the same exact-path and every-root import
requirements used for the GABA network.

**Figure G. Literature-supported Case-2 pathways ending in butyrate.** The
network is restricted to relationships with qualifying literature evidence for
both the donor-to-exact-precursor and recipient-to-butyrate reactions. Colours,
node roles, `n/N` edge labels, line-width thresholds, and snow-grey multi-precursor
input boxes are defined as in Figure F. Edge colour identifies the originating
analysis set; it does not indicate pathway direction or effect size. The
displayed connections remain GEM-supported structural potential rather than
experimentally confirmed exchange between the modeled organisms.

#### Complete relationship tables

The pathway figures are intentionally restricted to Case 2. The five Case-1
relationships and all 624 relationships lacking qualifying donor-to-precursor
support remain available in separate expandable tables below. Recipient-only
evidence is retained in the last table but is not treated as cross-feeding
literature support.

```{r complete-pathway-tables, results='asis'}
cat(as.character(make_expandable_table(
  "Open the complete 20-row Case-2 table: both constituent reactions supported",
  case2_pathway_table_view(case2_pathway_relationships)
)))
cat(as.character(make_expandable_table(
  "Open the complete 5-row Case-1 table: donor-to-precursor support only",
  pathway_table_view(case1_pathway_relationships)
)))
cat(as.character(make_expandable_table(
  "Open the complete 624-row table: no qualifying donor-to-precursor support",
  pathway_table_view(no_donor_support_pathway_relationships)
)))
```

The following optional export code is deliberately inactive. It can be enabled
later if separate supplementary CSV files are approved.

```{r optional-pathway-csv-export, eval=FALSE, include=FALSE}
# write.csv(
#   case2_pathway_table_view(case2_pathway_relationships),
#   "case2_both_constituent_reactions_supported.csv", row.names = FALSE
# )
# write.csv(
#   pathway_table_view(case1_pathway_relationships),
#   "case1_donor_to_precursor_supported_only.csv", row.names = FALSE
# )
# write.csv(
#   pathway_table_view(no_donor_support_pathway_relationships),
#   "pathway_relationships_no_qualifying_donor_support.csv",
#   row.names = FALSE
# )
```

### Primary literature used for constituent-reaction support

```{r literature-table, results='asis'}
literature <- data.frame(
  `Evidence side` = c(
    "Donor → precursor", "Donor → precursor", "Donor → precursor",
    "Donor → precursor", "Donor → precursor", "Recipient → target",
    "Recipient → target", "Recipient → target", "Recipient → target",
    "Recipient → target"
  ),
  `Taxon represented` = c(
    "Enterocloster bolteae / E. clostridioformis",
    "Ruminococcus bromii",
    "Alistipes onderdonkii",
    "Bacteroides ovatus",
    "Bacteroides xylanisolvens",
    "Agathobacter rectalis (formerly Eubacterium rectale)",
    "Agathobacter rectalis (formerly Eubacterium rectale)",
    "Bacteroides ovatus",
    "Gemmiger formicilis",
    "Roseburia hominis"
  ),
  `Literature strain(s)` = c(
    "WAL-14578; 2_1_49FAA; WAL-7855",
    "L2-63",
    "WAL 8169T / ATCC BAA-1178T",
    "Defined five-species culture including B. ovatus",
    "XB1AT / DSM 18836T",
    "ATCC 33656",
    "KLE1922",
    "DSM 1896",
    "ATCC 27749",
    "A2-183 / DSM 16839"
  ),
  `Supported reaction or metabolite` = c(
    "Extracellular glutamate and/or fumarate",
    "Ethanol fermentation product",
    "Succinate fermentation end product",
    "Succinate production in a defined coculture",
    "Succinate fermentation product",
    "Butyrate production",
    "GABA production",
    "GABA production",
    "Butyrate production",
    "Butyrate production"
  ),
  Citation = c(
    "Han et al. (2021), Nature",
    "Laverde Gomez et al. (2019), Environmental Microbiology",
    "Song et al. (2006), IJSEM",
    "Liu et al. (2020), Applied and Environmental Microbiology",
    "Chassard et al. (2008), IJSEM",
    "Rivière et al. (2015), Applied and Environmental Microbiology",
    "Strandwitz et al. (2019), Nature Microbiology",
    "Otaru et al. (2021), Frontiers in Microbiology",
    "Gossling and Moore (1975), International Journal of Systematic Bacteriology",
    "Bhattacharya et al. (2022), Microorganisms"
  ),
  DOI = c(
    "10.1038/s41586-021-03707-9",
    "10.1111/1462-2920.14454",
    "10.1099/ijs.0.64318-0",
    "10.1128/AEM.01905-19",
    "10.1099/ijs.0.65504-0",
    "10.1128/AEM.02089-15",
    "10.1038/s41564-018-0307-3",
    "10.3389/fmicb.2021.656895",
    "10.1099/00207713-25-2-202",
    "10.3390/microorganisms10122496"
  ),
  `Evidence scope and caveat` = c(
    "Direct cultured-strain extracellular metabolomics; the signal was condition- and strain-dependent.",
    "Exact-strain monoculture evidence; supports ethanol release, not the complete modeled route.",
    "Type-strain culture phenotype; supports succinate production at species level.",
    "Direct metabolite measurement in a defined coculture; attribution is less direct than isolated monoculture.",
    "Type-strain culture phenotype; supports succinate production at species level.",
    "Exact-strain cross-feeding experiment; supports butyrate production by the recipient species.",
    "Experimental strain panel; GABA production was pH-dependent and should not be generalized to every strain.",
    "Direct culture-supernatant quantification; production magnitude was strain- and condition-dependent.",
    "Exact type-strain physiological characterization; supports butyrate as a fermentation product.",
    "Exact-strain culture study; supports butyrate production during oligosaccharide utilization."
  ),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

literature$DOI <- sprintf(
  '<a href="https://doi.org/%s" target="_blank">%s</a>',
  literature$DOI,
  literature$DOI
)

kable(
  literature,
  format = "html",
  escape = FALSE,
  align = c("l", "l", "l", "l", "l", "l", "l"),
  caption = "Primary studies used to support at least one constituent reaction"
) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = TRUE,
    font_size = 11
  ) |>
  scroll_box(width = "100%", height = "620px")
```

#### Evidence statement for manuscript use

The literature layer should be reported as **constituent-reaction support**.
Case 1 supports only the donor-to-precursor reaction. Case 2 supports both terminal
reactions, but still does not experimentally confirm transfer of that precursor
between the modeled organisms. The GEM results remain predictions of structural
metabolic potential requiring experimental validation.

## Reproducibility

This report generates the plots and tables directly during knitting. It does not write separate figure or CSV files. The plotting code is retained in this R Markdown source but hidden in the rendered HTML.

```{r session-info}
sessionInfo()
```
