---
title: "CANNEX cross-feeding visualization reproducibility"
subtitle: "Validation Cases 01-08 and combined Core-6/Core-15 networks"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float: true
    code_folding: hide
    theme: flatly
params:
  data_root: null
  validation_root: null
  strict_input_hash: true
---

```{r reproduction-setup, include=FALSE}
if (!requireNamespace("knitr", quietly = TRUE)) {
  stop("The R package 'knitr' is required.")
}

current_input <- tryCatch(knitr::current_input(), error = function(e) "")
if (nzchar(current_input) && file.exists(current_input)) {
  visualization_directory <- dirname(normalizePath(
    current_input, winslash = "/", mustWork = TRUE
  ))
} else {
  visualization_directory <- normalizePath(
    getwd(), winslash = "/", mustWork = TRUE
  )
}

repository_root <- normalizePath(
  file.path(visualization_directory, ".."),
  winslash = "/", mustWork = TRUE
)
manuscript_source <- file.path(
  visualization_directory,
  "source",
  "CANNEX_CORE6_CORE15_CROSSFEEDING_SUMMARY.Rmd"
)

if (!file.exists(manuscript_source)) {
  stop("Missing locked manuscript Rmd: ", manuscript_source)
}

read_manuscript_chunk <- function(label) {
  lines <- readLines(manuscript_source, warn = FALSE)
  pattern <- paste0("^```\\{r[[:space:]]+", label, "([,}])")
  start <- grep(pattern, lines)
  if (length(start) != 1L) {
    stop("Expected exactly one manuscript chunk named '", label, "'.")
  }
  relative_end <- which(lines[(start + 1L):length(lines)] == "```")[1]
  if (is.na(relative_end)) {
    stop("The manuscript chunk '", label, "' is not closed.")
  }
  end <- start + relative_end
  paste(lines[(start + 1L):(end - 1L)], collapse = "\n")
}

evaluate_manuscript_chunk <- function(label) {
  expressions <- parse(text = read_manuscript_chunk(label))
  environment <- knitr::knit_global()
  value <- NULL
  for (expression in expressions) {
    value <- eval(expression, envir = environment)
  }
  value
}

# Load the exact packages, shared theme, and plotting helpers from the active
# manuscript source before supplying repository-relative input locations.
evaluate_manuscript_chunk("setup")

data_root <- if (is.null(params$data_root) || !nzchar(params$data_root)) {
  file.path(visualization_directory, "inputs", "gem_screening")
} else {
  params$data_root
}
validation_root <- if (
  is.null(params$validation_root) || !nzchar(params$validation_root)
) {
  file.path(repository_root, "validation")
} else {
  params$validation_root
}

crossfeeding_input_paths <- list(
  data_root = normalizePath(data_root, winslash = "/", mustWork = TRUE),
  validation_root = normalizePath(
    validation_root, winslash = "/", mustWork = TRUE
  )
)
crossfeeding_strict_input_hash <- isTRUE(params$strict_input_hash)
evaluate_manuscript_chunk("validation-data-preparation")
evaluate_manuscript_chunk("validation-control-parameters")
```

## Pipeline validation

Eight literature-defined cases were used to test whether the workflow retains
expected relationships and excludes expected null relationships. Cases 01 and
02 are matched triad controls: each contains a positive and a negative branch
while holding either the recipient or donor constant. Cases 03--05 are
independent positive anchors. Case 06 is a partial pathway match, whereas Cases
07--08 are informative non-passes in which target-production potential was
recovered but the literature-described donor edge was not reproduced.

```{r validation-outcome-plot, fig.width=10, fig.height=3.5, out.width='100%'}
evaluate_manuscript_chunk("validation-outcome-plot")
```

Cases 01 and 02 were matched triad controls expected to retain a
literature-supported positive branch while excluding a biologically
unsupported donor or recipient branch; the workflow reproduced both outcomes.
Cases 03--05 were independent positive controls representing
succinate-to-propionate, acetate-to-butyrate, and
lactate-to-hydrogen-sulfide cross-feeding, and all three expected relationships
were recovered. Case 06 was expected to recover the literature-described
*S. gordonii*--ornithine--*F. nucleatum* pathway, but the ornithine donor edge
was not retained. Instead, the GEMs supported a potential alternative
*S. gordonii*--L-arginine--*F. nucleatum* route to putrescine, resulting in a
partial match. Cases 07 and 08 recovered recipient-side
indole-3-propionate and deoxycholate production potential, respectively, but
did not recover the literature-described indole-3-lactate or cholate donor
edges and were classified as informative non-passes. Overall, five cases
passed formally, one produced a partial pathway match, and two exposed
donor-side limitations in the available GEMs or workflow.

**Figure V1. Pipeline-validation outcomes.** Bars give the number of the eight
literature-defined cases classified as a formal pass, partial pathway match, or
informative non-pass. Cases 01 and 02 are counted once each even though each
contains both a positive and negative branch. Colours indicate outcome class
only and do not quantify biological effect size.

The pathway panels use four edge meanings: deep-sky-blue solid arrows are
expected relationships retained by the workflow; black dashed T-bars are
expected null relationships correctly excluded; medium-orchid arrows are
recovered partial or recipient-side results; and violet-red dashed T-bars are
literature-expected edges that were not recovered. Grey dashed arrows provide
literature-route context only.

### Validation cases 01--02

```{r validation-cases-01-02}
evaluate_manuscript_chunk("validation-cases-01-02")
```

**Figure V2. Matched triad validation controls.** Case 01 holds the recipient
constant and contrasts a retained *Bifidobacterium adolescentis* L2-32
L-lactate donor with an excluded *Paraprevotella xylaniphila* YIT 11841 donor.
Case 02 holds the donor constant and contrasts retained L-lactate use by
*Eubacterium hallii* L2-7 with the excluded negative recipient
*Phascolarctobacterium succinatutens* YIT 12067. A dashed T-bar denotes the
prespecified absent relationship and therefore contributes to a formal pass.

### Validation cases 03--04

```{r validation-cases-03-04}
evaluate_manuscript_chunk("validation-cases-03-04")
```

**Figure V3. Independent short-chain-fatty-acid validation anchors.** Case 03
shows the retained *P. xylaniphila* YIT 11841 -> succinate ->
*P. succinatutens* YIT 12067 -> propionate relationship. Case 04 shows the
retained *Bifidobacterium longum* NCC2705 -> acetate -> *Eubacterium rectale*
ATCC 33656 -> butyrate relationship. Solid arrows denote retained expected
components.

### Validation case 05

```{r validation-case-05}
evaluate_manuscript_chunk("validation-case-05")
```

**Figure V4. Hydrogen-sulfide validation anchor.** Case 05 retained the
expected *B. adolescentis* L2-32 -> L-lactate -> *Desulfovibrio piger* ATCC
29098 -> hydrogen-sulfide relationship.

### Validation cases 06--08

```{r validation-cases-06-08}
evaluate_manuscript_chunk("validation-cases-06-08")
```

**Figure V5. Partial and informative non-pass validation cases.** Case 06
recovered putrescine-production potential and a GEM-supported potential
alternative *S. gordonii* CH1 -> L-arginine -> *F. nucleatum* ATCC 25586 route,
but did not recover the literature-described *S. gordonii* CH1 -> ornithine
donor edge. Case 07 recovered indole-3-propionate-production potential in
*Clostridium sporogenes* ATCC 15579, but the donor GEM did not represent the
expected indole-3-lactate metabolite. Case 08 recovered
deoxycholate-production potential and identified cholate on the recipient
side, but cholate production was not retained for *Bacteroides
thetaiotaomicron* VPI-5482. Violet-red dashed T-bars mark the expected donor
components that were not recovered.

```{r pathway-data-preparation, include=FALSE}
evaluate_manuscript_chunk("exact-count-table")
evaluate_manuscript_chunk("sankey-data")
evaluate_manuscript_chunk("pathway-data-preparation")
```

## Combined GABA network

This figure contains nine unique Case-2 pathways after combining the two
analysis sets. No complete GABA pathway was duplicated between Core-6 and
Core-15 after normalizing *Eubacterium rectale* to *Agathobacter rectalis* for
display.

```{r gaba-pathway-parameters, include=FALSE}
evaluate_manuscript_chunk("gaba-pathway-parameters")
```

```{r gaba-pathway-plot}
evaluate_manuscript_chunk("gaba-pathway-plot")
```

The GABA network contained four Core-6 Case-2 relationships (two
single-precursor and two multi-precursor) and five Core-15 Case-2 relationships
(all multi-precursor). No complete relationship was duplicated between the two
sets after taxonomic display-name normalization. For every displayed
multi-precursor route, at least one exact-path recipient GEM could import and
transport every metabolite in the displayed input set.

**Figure F. Literature-supported Case-2 pathways ending in GABA.** The network
shows only relationships for which both donor-to-exact-precursor and
recipient-to-GABA evidence met the prespecified literature rule.
Light-sky-blue edges denote Core-6, steel-blue edges denote Core-15, and
slate-blue edges would denote a relationship present in both sets. Node colours
distinguish potential donors (deep sky blue), focal precursors (thistle),
potential recipients (slate blue), and GABA (pale violet red). A snow-grey
intermediate box lists the complete input combination for a displayed
multi-precursor path. Edge labels report supporting/tested GEMs (`n/N`); line
width denotes low (<=30%), intermediate (>30% to <=70%), or high (>70%)
strain-panel support. These are structural-potential relationships, not direct
evidence of in vivo metabolite transfer.

## Combined butyrate network

This figure contains eleven unique Case-2 pathways after combining the two
analysis sets. No complete butyrate pathway was duplicated between Core-6 and
Core-15 after taxonomic display-name normalization.

```{r butyrate-pathway-parameters, include=FALSE}
evaluate_manuscript_chunk("butyrate-pathway-parameters")
```

```{r butyrate-pathway-plot}
evaluate_manuscript_chunk("butyrate-pathway-plot")
```

The butyrate network contained five Core-6 Case-2 relationships (three
single-precursor and two multi-precursor) and six Core-15 Case-2 relationships
(all single-precursor). No complete relationship was duplicated between the
two sets after taxonomic display-name normalization. Each displayed
multi-precursor input set passed the same exact-path and every-root import
requirements used for the GABA network.

**Figure G. Literature-supported Case-2 pathways ending in butyrate.** The
network is restricted to relationships with qualifying literature evidence for
both the donor-to-exact-precursor and recipient-to-butyrate reactions. Colours,
node roles, `n/N` edge labels, line-width thresholds, and snow-grey
multi-precursor input boxes are defined as in Figure F. Edge colour identifies
the originating analysis set; it does not indicate pathway direction or effect
size. The displayed connections remain GEM-supported structural potential
rather than experimentally confirmed exchange between the modeled organisms.

## Reproducibility information

The source manuscript code checks the locked validation and pathway-input
SHA-256 values before plotting. The software versions used for this rendering
are recorded below.

```{r session-info}
sessionInfo()
```
